### 2024-10-27 paper_v01/article.tex
first commit

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{siunitx}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation (and optionally ORCID link)
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Fernando Ferraz Ribeiro~\orcidlink{0000-0002-0685-4774}\\Universidade Federal da Bahia
   \And Gilney Figueira Zebende~\orcidlink{0000-0003-2420-9805}\\Universidade Estadual\\de Feira de Santana}
\Plainauthor{Fernando Ferraz Ribeiro, Gilney Figueira Zebende}
%  and $DMC_x^2$
%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$ methods}
\Plaintitle{A Python/Zig optimized and customizable implementation for the Pdcca and DMCx2 methods}
\Shorttitle{\proglang{Python}/\proglang{Zig} implementation for the $\rho_{DCCA}$ and $DMC_x^2$}

%% - \Abstract{} almost as usual
\Abstract{
  This short article illustrates how to write a manuscript for the
  \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
  Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
  it is recommended to keep the {\LaTeX} code as simple as possible,
  i.e., avoid inclusion of packages/commands that are not necessary.
  For outlining the typical structure of a JSS article some brief text snippets
  are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
  discussing count data regression in \proglang{R}. Editorial comments and
  instructions are marked by vertical bars.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{$\rho_{DCCA}$, $DMC_x^2$, optimization, \proglang{Python}, \proglang{Zig}}
\Plainkeywords{Pdcca, DMCx2, optimization, Python, Zig}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Fernando Ferraz Ribeiro\\
  Universidade Federal da Bahia\\
  Faculty of Achitecture\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{fernando.ribeiro@ufba.br}\\
  \emph{and}\\
  Centro Universitário Senai-Cimatec\\

  % URL: \url{https://www.zeileis.org/}
}

\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section{introduction} \label{sec:intro}

The $\rho_{DCCA}$~\citep{Zebende2011} is a widely used coefficient that measures the cross-correlation between tow non-stationary time series. It´s an extension of the Detrended Fluctuation Analysis ($DFA$)~\citep{Peng_1994} and the Detrended Cross-correlation Analysis ($DCCA$)~\citep{Podobnik2008}: while the $DFA$ calculates the self-affinity and long-memory properties of a time series data, and the $DCCA$ analyses power-law cross correlations between two different non-stationarity time series, the $\rho_{DCCA}$ coefficient quantifies this cross-correlation in simple values ranging from $-1$ to $1$, where $-1$ indicates a perfect anti-correlation between the series, $1$ a perfect correlation and zero ($0$) no correlation at all.

The Detrended Multiple Cross-Correlation Coefficient \citep{Zebende2018} ($DMC_x^2$) is a generalization of the $\rho_{DCCA}$ coefficient that correlates one time series (dependent variable) a number of time series (independent variables). The $DMC_x^2$ values ranges from zero ($0$), indicating no correlation to $1$, meaning perfect correlation or anti-correlation between the dependent and the independent variables.

This paper presents the \pkg{Zebende} \proglang{Python} package, an implementation of the $DFA$, $DCCA$, $\rho_{DCCCA}$, $DMC_x^2$ and utility functions related to the methods. In section \ref{sec:calculations} the steps for calculating the $\rho_{DCCCA}$ and $DMC_x^2$ are presented and discussed. Section \ref{sec:optimization} shows how this library was implemented, the optimization technics and the recommended steps to use the library. In Section \ref{sec:results} the \pkg{Zebende} package is compared with other packages for \proglang{Python} and \proglang{R} that calculates the $\rho_{DCCA}$ in terms of performance and usability, leading to the conclusions in \ref{sec:summary}.

\section{Algorithms of the coefficients}\label{sec:calculations}

\begin{equation}
  {DMC}_{x}^{2}  \equiv \rho_{Y,X_{i}}(n)^{T} \times \rho^{-1}(n) \times \rho_{Y,X_{i}}(n)
  \label{eq:dmcx2}
\end{equation}

The ${DMC}_{x}^{2}$ ranges from zero $(0)$, where the time-series do not correlate, to one $(1)$ meaning that the time-series have either a perfect correlation or a perfect anti-correlation. In this equation, $\rho^{-1}(n)$ represent the inverse matrix of all possible combinations of $\rho_{DCCA}$ between the independent variables, in other words:

\begin{equation}
  \rho^{-1}(n) = \left(\begin{matrix}
    1                     & \rho_{X_{1},X_{2}}(n) & \rho_{X_{1},X_{3}}(n) & \dots & \rho_{X_{1},X_{j}}(n) \\
    \rho_{X_{2},X_{1}}(n) & 1                     & \rho_{X_{2},X_{3}}(n) & \dots & \rho_{X_{2},X_{j}}(n) \\
    \vdots                & \vdots                & \vdots                & \dots & \vdots                \\
    \rho_{X_{j},X_{1}}(n) & \rho_{X_{j},X_{2}}(n) & \rho_{X_{j},X_{3}}(n) & \dots & 1                     \\
  \end{matrix}\right)^{-1}
  \label{eq:p_dcca_matrix}
\end{equation}

\begin{equation}
  \rho_{Y,X_k}(n)^T=[\rho_{Y,X_1}(n), \rho_{Y,X_2}(n),\cdots,\rho_{Y,X_j}(n)]
  \label{eq:rhocol}
\end{equation}

\begin{equation}
  {\rho}_{DCCA}(n) = \frac{F_{DCCA~(x\alpha,~x\beta)}^{2}(n)}
  {F_{DFA~(x\alpha)}(n) \times F_{DFA~(x\beta)}(n)}
  \label{eq:p_dcca}
\end{equation}

\begin{description}

  \label{steps:dfa}
  \item Taking a time series \(\{x_{i}\}\) with \(i\) ranging from \(1\) to \(N\), the integrated series \(X_{k}\) is calculated by \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) with \(k\) also ranging from \(1\) to \(N\);
  \item the  \(X_{k}\) series is divided in \(N - n\) boxes of size \(n\)(time scale),each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting \(\widetilde{X}_{k, i}\) with \( i \le k \le (i + n) \) (detrended values);
  \item in each box is calculated: \(f_{DFA}^{2}(n, i) = \frac{1}{1+n} \sum_{k=i}^{i + n}(X_{k}-\widetilde{X}_{k, i})^{2}\)
  \item for all the boxes of a time scale, the DFA is calculated as:\\[10pt]
        \(F_{DFA}(n) = \sqrt{\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DFA}^{2}(n, i)}\);
  \item for a number of different timescales (n), with possible values \( 4 \le n \le \frac{N}{4}\) the \(F_{DFA}\) function is calculated to find a relation among \(F_{DFA} \times n\)

\end{description}

\begin{description}
  \label{steps:DCCA}
  \item Taking two time series with the same length \(\{x\alpha_{i}\}\) and \(\{x\beta_{i}\}\) with \(i\) ranging from \(1\) to \(N\),
        the integrated series \(X\alpha_{k}\) and \(X\beta_{k}\) are calculated by
        \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) for each series, with \(k\) also ranging from \(1\) to \(N\);
  \item \(X\alpha_{k}\) and \(X\beta_{k}\) series are divided in \(N - n\) boxes of size \(n\)(time scale),
        each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting
        \(\widetilde{X\alpha}_{k, i}\) and \(\widetilde{X\beta}_{k, i}\),
        for series \(\{x\alpha_{i}\}\) and \(\{x\beta_{i}\}\) respectively,
        with \( i \le k \le (i + n) \) (detrended values);
  \item in each box is calculated: \(f_{DCCA}^{2}(n, i) =
        \frac{1}{1+n} \sum_{k=i}^{i + n}(X\alpha_{k}-\widetilde{X\alpha}_{k, i}) \times (X\beta_{k}-\widetilde{X\beta}_{k, i})\)
  \item for all the boxes of a time scale, the DFA is calculated as:\\[10pt]
        \(F_{DCCA}^{2}(n) =\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DCCA}^{2}(n, i)\);
  \item for a number of different timescales (n), with possible values
        \( 4 \le n \le \frac{N}{4}\) the \(F_{DCCA}^{2}\) function is calculated to find a relation among
        \(F_{DCCA}^{2} \times n\)
\end{description}

Modeling count variables is a common task in economics and the social sciences.
The classical Poisson regression model for count data is often of limited use in
these disciplines because empirical count data sets typically exhibit
overdispersion and/or an excess number of zeros. The former issue can be
addressed by extending  the plain Poisson regression model in various
directions: e.g., using sandwich covariances or estimating an additional
dispersion parameter (in a so-called quasi-Poisson model). Another more formal
way is to use a negative binomial (NB) regression. All of these models belong to
the family of generalized linear models (GLMs). However, although these models
typically can capture overdispersion rather well, they are in many applications
not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
increased interest in zero-augmented models that address this issue by a second
model component capturing zero counts. An overview of count data models in
econometrics, including  hurdle and zero-inflated models, is provided in
\cite{Cameron+Trivedi:2013}.

In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
diagnostics and inference. The manuscript that this document is based on
\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
package \citep{Jackman:2015}. Of course, much more software could be discussed
here, including (but not limited to) generalized additive models for count data
as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006},
\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Zebende package: implementation and optimization} \label{sec:optimization}

The basic Poisson regression model for count data is a special case of the GLM
framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
log link and a linear predictor
%
\begin{equation} \label{eq:mean}
  \log(\mu_i) \quad = \quad x_i^\top \beta,
\end{equation}
%
where the regression coefficients $\beta$ are estimated by maximum likelihood
(ML) using the iterative weighted least squares (IWLS) algorithm.

\begin{leftbar}
  Note that around the \verb|{equation}| above there should be no spaces (avoided
  in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
  not a new paragraph started.
\end{leftbar}

\proglang{R} provides a very flexible implementation of the general GLM
framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
\pkg{stats} package. Its most important arguments are
\begin{Code}
  glm(formula, data, subset, na.action, weights, offset,
  family = gaussian, start = NULL, control = glm.control(...),
  model = TRUE, y = TRUE, x = FALSE, ...)
\end{Code}
where \code{formula} plus \code{data} is the now standard way of specifying
regression relationships in \proglang{R}/\proglang{S} introduced in
\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
standard  for setting up formula-based regression models in
\proglang{R}/\proglang{S}. The arguments in the second line control aspects
specific to GLMs while the arguments in the last line specify which components
are returned in the fitted model object (of class \class{glm} which inherits
from \class{lm}). For further arguments to \fct{glm} (including alternative
specifications of starting values) see \code{?glm}. For estimating a Poisson
model \code{family = poisson} has to be specified.

\begin{leftbar}
  As the synopsis above is a code listing that is not meant to be executed,
  one can use either the dedicated \verb|{Code}| environment or a simple
  \verb|{verbatim}| environment for this. Again, spaces before and after should be
  avoided.

  Finally, there might be a reference to a \verb|{table}| such as
  Table~\ref{tab:overview}. Usually, these are placed at the top of the page
  (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
  column headers and captions in sentence style, and if possible avoiding vertical
  lines.
\end{leftbar}

\begin{table}[t!]
  \centering
  \begin{tabular}{lllp{7.4cm}}
    \hline
    Type           & Distribution & Method   & Description                        \\ \hline
    GLM            & Poisson      & ML       & Poisson regression: classical GLM,
    estimated by maximum likelihood (ML)                                          \\
                   &              & Quasi    & ``Quasi-Poisson regression'':
    same mean function, estimated by
    quasi-ML (QML) or equivalently
    generalized estimating equations (GEE),
    inference adjustment via estimated
    dispersion parameter                                                          \\
                   &              & Adjusted & ``Adjusted Poisson regression'':
    same mean function, estimated by
    QML/GEE, inference adjustment via
    sandwich covariances                                                          \\
                   & NB           & ML       & NB regression: extended GLM,
    estimated by ML including additional
    shape parameter                                                               \\ \hline
    Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
    hurdle Poisson                                                                \\
                   & NB           & ML       & Zero-inflated NB (ZINB),
    hurdle NB                                                                     \\ \hline
  \end{tabular}
  \caption{\label{tab:overview} Overview of various count regression models. The
    table is usually placed at the top of the page (\texttt{[t!]}), centered
    (\texttt{centering}), has a caption below the table, column headers and captions
    are in sentence style, and if possible vertical lines should be avoided.}
\end{table}

%% -- Illustrations ------------------------------------------------------------

%% - Virtually all JSS manuscripts list source code along with the generated
%%   output. The style files provide dedicated environments for this.
%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%%   or knitr using the render_sweave() hook - are used (without the need to
%%   load Sweave.sty).
%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
%% - The code input should use "the usual" command prompt in the respective
%%   software system.
%% - For R code, the prompt "R> " should be used with "+  " as the
%%   continuation prompt.
%% - Comments within the code chunks should be avoided - these should be made
%%   within the regular LaTeX text.

\section{Results} \label{sec:results}

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> data("quine", package = "MASS")
  \end{CodeInput}
\end{CodeChunk}
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

\begin{leftbar}
  For code input and output, the style files provide dedicated environments.
  Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
  or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
  produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
  hook. Please make sure that all code is properly spaced, e.g., using
  \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
  use ``the usual'' command prompt in the respective software system. For
  \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
  the continuation prompt. Generally, comments within the code chunks should be
  avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
  before and after code input/output should be avoided (see above).
\end{leftbar}

\begin{figure}[t!]
  \centering
  \includegraphics{article-visualization}
  \caption{\label{fig:quine} Frequency distribution for number of days absent
    from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
    +    family = poisson)
  \end{CodeInput}
\end{CodeChunk}
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> library("MASS")
    R> m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
  \end{CodeInput}
\end{CodeChunk}
%
In a comparison with the BIC the latter model is clearly preferred.
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> BIC(m_pois, m_nbin)
  \end{CodeInput}
  \begin{CodeOutput}
    df      BIC
    m_pois 18 2046.851
    m_nbin 19 1157.235
  \end{CodeOutput}
\end{CodeChunk}
%
Hence, the full summary of that model is shown below.
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> summary(m_nbin)
  \end{CodeInput}
  \begin{CodeOutput}
    Call:
    glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
    init.theta = 1.60364105, link = log)

    Deviance Residuals:
    Min       1Q   Median       3Q      Max
    -3.0857  -0.8306  -0.2620   0.4282   2.0898

    Coefficients: (1 not defined because of singularities)
    Estimate Std. Error z value Pr(>|z|)
    (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
    EthN        -0.24591    0.39135  -0.628  0.52977
    SexM        -0.77181    0.38021  -2.030  0.04236 *
    AgeF1       -0.02546    0.41615  -0.061  0.95121
    AgeF2       -0.54884    0.54393  -1.009  0.31296
    AgeF3       -0.25735    0.40558  -0.635  0.52574
    LrnSL        0.38919    0.48421   0.804  0.42153
    EthN:SexM    0.36240    0.29430   1.231  0.21818
    EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876
    EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 **
    EthN:AgeF3   0.04721    0.44883   0.105  0.91622
    EthN:LrnSL   0.06847    0.34040   0.201  0.84059
    SexM:AgeF1   0.02257    0.47360   0.048  0.96198
    SexM:AgeF2   1.55330    0.51325   3.026  0.00247 **
    SexM:AgeF3   1.25227    0.45539   2.750  0.00596 **
    SexM:LrnSL   0.07187    0.40805   0.176  0.86019
    AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870
    AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363
    AgeF3:LrnSL       NA         NA      NA       NA
    ---
    Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

    (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)

    Null deviance: 235.23  on 145  degrees of freedom
    Residual deviance: 167.53  on 128  degrees of freedom
    AIC: 1100.5

    Number of Fisher Scoring iterations: 1

    Theta:  1.604
    Std. Err.:  0.214

    2 x log-likelihood:  -1062.546
  \end{CodeOutput}
\end{CodeChunk}

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

\begin{leftbar}
  As usual \dots
\end{leftbar}

%% -- Optional special unnumbered sections -------------------------------------

\section*{Computational details}

\begin{leftbar}
  If necessary or useful, information about certain computational details
  such as version numbers, operating systems, or compilers could be included
  in an unnumbered section. Also, auxiliary packages (say, for visualizations,
  maps, tables, \dots) that are not cited in the main text can be credited here.
\end{leftbar}

The results in this paper were obtained using
\proglang{R}~3.4.1 with the
\pkg{MASS}~7.3.47 package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

\begin{leftbar}
  All acknowledgments (note the AE spelling) should be collected in this
  unnumbered section before the references. It may contain the usual information
  about funding and feedback from colleagues/reviewers/etc. Furthermore,
  information such as relative contributions of the authors may be added here
  (if any).
\end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

\begin{appendix}

  \section{More technical details} \label{app:technical}

  \begin{leftbar}
    Appendices can be included after the bibliography (with a page break). Each
    section within the appendix should have a proper section title (rather than
    just \emph{Appendix}).

    For more technical style details, please check out JSS's style FAQ at
    \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
    which includes the following topics:
    \begin{itemize}
      \item Title vs.\ sentence case.
      \item Graphics formatting.
      \item Naming conventions.
      \item Turning JSS manuscripts into \proglang{R} package vignettes.
      \item Trouble shooting.
      \item Many other potentially helpful details\dots
    \end{itemize}
  \end{leftbar}

  \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

  \begin{leftbar}
    References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
    references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
    \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
    formats of author-year citations and allow to include additional details (e.g.,
    pages, chapters, \dots) in brackets. In case you are not familiar with these
    commands see the JSS style FAQ for details.

    Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
    when acquiring the entries automatically from mixed online sources. However,
    it is important that informations are complete and presented in a consistent
    style to avoid confusions. JSS requires the following format.
    \begin{itemize}
      \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
            be used in the references.
      \item Titles should be in title case.
      \item Journal titles should not be abbreviated and in title case.
      \item DOIs should be included where available.
      \item Software should be properly cited as well. For \proglang{R} packages
            \code{citation("pkgname")} typically provides a good starting point.
    \end{itemize}
  \end{leftbar}

\end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2024-10-28 paper_v01/article.tex
update paper

@@
 \Abstract{
-  This short article illustrates how to write a manuscript for the
-  \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
-  Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
-  it is recommended to keep the {\LaTeX} code as simple as possible,
-  i.e., avoid inclusion of packages/commands that are not necessary.
-  For outlining the typical structure of a JSS article some brief text snippets
-  are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
-  discussing count data regression in \proglang{R}. Editorial comments and
-  instructions are marked by vertical bars.
+  This paper presents tha \pkg{Zebende}, a \proglang{Python} package writen in \proglang{Python} and \proglang{Zig}, that calculates the $DFA$, $DCCA$ $\rho_{DCCA}$ and the $DMC_x^2$.
+The package presents an optimized algorithm that significantly improves the calculations speed.
+A comparison with ohter packages that calculates the .The package is also the firts to implement the $DMC_x^2$ corefficient for any number of series.
 }
@@
 
-The ${DMC}_{x}^{2}$ ranges from zero $(0)$, where the time-series do not correlate, to one $(1)$ meaning that the time-series have either a perfect correlation or a perfect anti-correlation.
-In this equation, $\rho^{-1}(n)$ represent the inverse matrix of all possible combinations of $\rho_{DCCA}$ between the independent variables, in other words:
-
 \begin{equation}
@@
 
-Modeling count variables is a common task in economics and the social sciences.
-The classical Poisson regression model for count data is often of limited use in
-these disciplines because empirical count data sets typically exhibit
-overdispersion and/or an excess number of zeros. The former issue can be
-addressed by extending  the plain Poisson regression model in various
-directions: e.g., using sandwich covariances or estimating an additional
-dispersion parameter (in a so-called quasi-Poisson model). Another more formal
-way is to use a negative binomial (NB) regression. All of these models belong to
-the family of generalized linear models (GLMs). However, although these models
-typically can capture overdispersion rather well, they are in many applications
-not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
-increased interest in zero-augmented models that address this issue by a second
-model component capturing zero counts. An overview of count data models in
-econometrics, including  hurdle and zero-inflated models, is provided in
-\cite{Cameron+Trivedi:2013}.
-
-In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
-\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
-\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
-diagnostics and inference. The manuscript that this document is based on
-\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
-count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
-package \citep{Jackman:2015}. Of course, much more software could be discussed
-here, including (but not limited to) generalized additive models for count data
-as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006},
-\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.
-
 %% -- Manuscript ---------------------------------------------------------------
@@
 \section{Zebende package: implementation and optimization} \label{sec:optimization}
-
-The basic Poisson regression model for count data is a special case of the GLM
-framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
-response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
-$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
-$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
-log link and a linear predictor
-%
-\begin{equation} \label{eq:mean}
-  \log(\mu_i) \quad = \quad x_i^\top \beta,
-\end{equation}
-%
-where the regression coefficients $\beta$ are estimated by maximum likelihood
-(ML) using the iterative weighted least squares (IWLS) algorithm.
-
-\begin{leftbar}
-  Note that around the \verb|{equation}| above there should be no spaces (avoided
-  in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
-  not a new paragraph started.
-\end{leftbar}
-
-\proglang{R} provides a very flexible implementation of the general GLM
-framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
-\pkg{stats} package. Its most important arguments are
-\begin{Code}
-  glm(formula, data, subset, na.action, weights, offset,
-  family = gaussian, start = NULL, control = glm.control(...),
-  model = TRUE, y = TRUE, x = FALSE, ...)
-\end{Code}
-where \code{formula} plus \code{data} is the now standard way of specifying
-regression relationships in \proglang{R}/\proglang{S} introduced in
-\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
-(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
-standard  for setting up formula-based regression models in
-\proglang{R}/\proglang{S}. The arguments in the second line control aspects
-specific to GLMs while the arguments in the last line specify which components
-are returned in the fitted model object (of class \class{glm} which inherits
-from \class{lm}). For further arguments to \fct{glm} (including alternative
-specifications of starting values) see \code{?glm}. For estimating a Poisson
-model \code{family = poisson} has to be specified.
-
-\begin{leftbar}
-  As the synopsis above is a code listing that is not meant to be executed,
-  one can use either the dedicated \verb|{Code}| environment or a simple
-  \verb|{verbatim}| environment for this. Again, spaces before and after should be
-  avoided.
-
-  Finally, there might be a reference to a \verb|{table}| such as
-  Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-  (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-  column headers and captions in sentence style, and if possible avoiding vertical
-  lines.
-\end{leftbar}
-
-\begin{table}[t!]
-  \centering
-  \begin{tabular}{lllp{7.4cm}}
-    \hline
-    Type           & Distribution & Method   & Description                        \\ \hline
-    GLM            & Poisson      & ML       & Poisson regression: classical GLM,
-    estimated by maximum likelihood (ML)                                          \\
-                   &              & Quasi    & ``Quasi-Poisson regression'':
-    same mean function, estimated by
-    quasi-ML (QML) or equivalently
-    generalized estimating equations (GEE),
-    inference adjustment via estimated
-    dispersion parameter                                                          \\
-                   &              & Adjusted & ``Adjusted Poisson regression'':
-    same mean function, estimated by
-    QML/GEE, inference adjustment via
-    sandwich covariances                                                          \\
-                   & NB           & ML       & NB regression: extended GLM,
-    estimated by ML including additional
-    shape parameter                                                               \\ \hline
-    Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
-    hurdle Poisson                                                                \\
-                   & NB           & ML       & Zero-inflated NB (ZINB),
-    hurdle NB                                                                     \\ \hline
-  \end{tabular}
-  \caption{\label{tab:overview} Overview of various count regression models. The
-    table is usually placed at the top of the page (\texttt{[t!]}), centered
-    (\texttt{centering}), has a caption below the table, column headers and captions
-    are in sentence style, and if possible vertical lines should be avoided.}
-\end{table}
 
@@
 
-For a simple illustration of basic Poisson and NB count regression the
-\code{quine} data from the \pkg{MASS} package is used. This provides the number
-of \code{Days} that children were absent from school in Australia in a
-particular year, along with several covariates that can be employed as regressors.
-The data can be loaded by
-%
-\begin{CodeChunk}
-  \begin{CodeInput}
-    R> data("quine", package = "MASS")
-  \end{CodeInput}
-\end{CodeChunk}
-%
-and a basic frequency distribution of the response variable is displayed in
-Figure~\ref{fig:quine}.
-
-\begin{leftbar}
-  For code input and output, the style files provide dedicated environments.
-  Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-  or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-  produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-  hook. Please make sure that all code is properly spaced, e.g., using
-  \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-  use ``the usual'' command prompt in the respective software system. For
-  \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-  the continuation prompt. Generally, comments within the code chunks should be
-  avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-  before and after code input/output should be avoided (see above).
-\end{leftbar}
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics{article-visualization}
-  \caption{\label{fig:quine} Frequency distribution for number of days absent
-    from school.}
-\end{figure}
-
-As a first model for the \code{quine} data, we fit the basic Poisson regression
-model. (Note that JSS prefers when the second line of code is indented by two
-spaces.)
-%
-\begin{CodeChunk}
-  \begin{CodeInput}
-    R> m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-    +    family = poisson)
-  \end{CodeInput}
-\end{CodeChunk}
-%
-To account for potential overdispersion we also consider a negative binomial
-GLM.
-%
-\begin{CodeChunk}
-  \begin{CodeInput}
-    R> library("MASS")
-    R> m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-  \end{CodeInput}
-\end{CodeChunk}
-%
-In a comparison with the BIC the latter model is clearly preferred.
-%
-\begin{CodeChunk}
-  \begin{CodeInput}
-    R> BIC(m_pois, m_nbin)
-  \end{CodeInput}
-  \begin{CodeOutput}
-    df      BIC
-    m_pois 18 2046.851
-    m_nbin 19 1157.235
-  \end{CodeOutput}
-\end{CodeChunk}
-%
-Hence, the full summary of that model is shown below.
-%
-\begin{CodeChunk}
-  \begin{CodeInput}
-    R> summary(m_nbin)
-  \end{CodeInput}
-  \begin{CodeOutput}
-    Call:
-    glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-    init.theta = 1.60364105, link = log)
-
-    Deviance Residuals:
-    Min       1Q   Median       3Q      Max
-    -3.0857  -0.8306  -0.2620   0.4282   2.0898
-
-    Coefficients: (1 not defined because of singularities)
-    Estimate Std. Error z value Pr(>|z|)
-    (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
-    EthN        -0.24591    0.39135  -0.628  0.52977
-    SexM        -0.77181    0.38021  -2.030  0.04236 *
-    AgeF1       -0.02546    0.41615  -0.061  0.95121
-    AgeF2       -0.54884    0.54393  -1.009  0.31296
-    AgeF3       -0.25735    0.40558  -0.635  0.52574
-    LrnSL        0.38919    0.48421   0.804  0.42153
-    EthN:SexM    0.36240    0.29430   1.231  0.21818
-    EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876
-    EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 **
-    EthN:AgeF3   0.04721    0.44883   0.105  0.91622
-    EthN:LrnSL   0.06847    0.34040   0.201  0.84059
-    SexM:AgeF1   0.02257    0.47360   0.048  0.96198
-    SexM:AgeF2   1.55330    0.51325   3.026  0.00247 **
-    SexM:AgeF3   1.25227    0.45539   2.750  0.00596 **
-    SexM:LrnSL   0.07187    0.40805   0.176  0.86019
-    AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870
-    AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363
-    AgeF3:LrnSL       NA         NA      NA       NA
-    ---
-    Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-
-    (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
-
-    Null deviance: 235.23  on 145  degrees of freedom
-    Residual deviance: 167.53  on 128  degrees of freedom
-    AIC: 1100.5
-
-    Number of Fisher Scoring iterations: 1
-
-    Theta:  1.604
-    Std. Err.:  0.214
-
-    2 x log-likelihood:  -1062.546
-  \end{CodeOutput}
-\end{CodeChunk}
-
 %% -- Summary/conclusions/discussion -------------------------------------------
@@
 
-\section*{Computational details}
+% \section*{Computational details}
 
-\begin{leftbar}
-  If necessary or useful, information about certain computational details
-  such as version numbers, operating systems, or compilers could be included
-  in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-  maps, tables, \dots) that are not cited in the main text can be credited here.
-\end{leftbar}
+% \begin{leftbar}
+%   If necessary or useful, information about certain computational details
+%   such as version numbers, operating systems, or compilers could be included
+%   in an unnumbered section. Also, auxiliary packages (say, for visualizations,
+%   maps, tables, \dots) that are not cited in the main text can be credited here.
+% \end{leftbar}
 
-The results in this paper were obtained using
-\proglang{R}~3.4.1 with the
-\pkg{MASS}~7.3.47 package. \proglang{R} itself
-and all packages used are available from the Comprehensive
-\proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/}.
+% The results in this paper were obtained using
+% \proglang{R}~3.4.1 with the
+% \pkg{MASS}~7.3.47 package. \proglang{R} itself
+% and all packages used are available from the Comprehensive
+% \proglang{R} Archive Network (CRAN) at
+% \url{https://CRAN.R-project.org/}.
 
-\section*{Acknowledgments}
+% \section*{Acknowledgments}
 
-\begin{leftbar}
-  All acknowledgments (note the AE spelling) should be collected in this
-  unnumbered section before the references. It may contain the usual information
-  about funding and feedback from colleagues/reviewers/etc. Furthermore,
-  information such as relative contributions of the authors may be added here
-  (if any).
-\end{leftbar}
+% \begin{leftbar}
+%   All acknowledgments (note the AE spelling) should be collected in this
+%   unnumbered section before the references. It may contain the usual information
+%   about funding and feedback from colleagues/reviewers/etc. Furthermore,
+%   information such as relative contributions of the authors may be added here
+%   (if any).
+% \end{leftbar}
 
@@
 
-\newpage
+% \newpage
 
-\begin{appendix}
+% \begin{appendix}
 
-  \section{More technical details} \label{app:technical}
+%   \section{More technical details} \label{app:technical}
 
-  \begin{leftbar}
-    Appendices can be included after the bibliography (with a page break). Each
-    section within the appendix should have a proper section title (rather than
-    just \emph{Appendix}).
+%   \begin{leftbar}
+%     Appendices can be included after the bibliography (with a page break). Each
+%     section within the appendix should have a proper section title (rather than
+%     just \emph{Appendix}).
 
-    For more technical style details, please check out JSS's style FAQ at
-    \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-    which includes the following topics:
-    \begin{itemize}
-      \item Title vs.\ sentence case.
-      \item Graphics formatting.
-      \item Naming conventions.
-      \item Turning JSS manuscripts into \proglang{R} package vignettes.
-      \item Trouble shooting.
-      \item Many other potentially helpful details\dots
-    \end{itemize}
-  \end{leftbar}
+%     For more technical style details, please check out JSS's style FAQ at
+%     \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
+%     which includes the following topics:
+%     \begin{itemize}
+%       \item Title vs.\ sentence case.
+%       \item Graphics formatting.
+%       \item Naming conventions.
+%       \item Turning JSS manuscripts into \proglang{R} package vignettes.
+%       \item Trouble shooting.
+%       \item Many other potentially helpful details\dots
+%     \end{itemize}
+%   \end{leftbar}
 
-  \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
+%   \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
 
-  \begin{leftbar}
-    References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-    references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-    \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-    formats of author-year citations and allow to include additional details (e.g.,
-    pages, chapters, \dots) in brackets. In case you are not familiar with these
-    commands see the JSS style FAQ for details.
+%   \begin{leftbar}
+%     References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
+%     references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
+%     \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
+%     formats of author-year citations and allow to include additional details (e.g.,
+%     pages, chapters, \dots) in brackets. In case you are not familiar with these
+%     commands see the JSS style FAQ for details.
 
-    Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-    when acquiring the entries automatically from mixed online sources. However,
-    it is important that informations are complete and presented in a consistent
-    style to avoid confusions. JSS requires the following format.
-    \begin{itemize}
-      \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-            be used in the references.
-      \item Titles should be in title case.
-      \item Journal titles should not be abbreviated and in title case.
-      \item DOIs should be included where available.
-      \item Software should be properly cited as well. For \proglang{R} packages
-            \code{citation("pkgname")} typically provides a good starting point.
-    \end{itemize}
-  \end{leftbar}
+%     Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
+%     when acquiring the entries automatically from mixed online sources. However,
+%     it is important that informations are complete and presented in a consistent
+%     style to avoid confusions. JSS requires the following format.
+%     \begin{itemize}
+%       \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
+%             be used in the references.
+%       \item Titles should be in title case.
+%       \item Journal titles should not be abbreviated and in title case.
+%       \item DOIs should be included where available.
+%       \item Software should be properly cited as well. For \proglang{R} packages
+%             \code{citation("pkgname")} typically provides a good starting point.
+%     \end{itemize}
+%   \end{leftbar}
 
-\end{appendix}
+% \end{appendix}
 

### 2024-11-01 paper_v01/article.tex
DMCx2 flowchart update

@@
 \Abstract{
-  This paper presents tha \pkg{Zebende}, a \proglang{Python} package writen in \proglang{Python} and \proglang{Zig}, that calculates the $DFA$, $DCCA$ $\rho_{DCCA}$ and the $DMC_x^2$.
+  This paper presents tha \pkg{Zebende}, a \proglang{Python} package written in \proglang{Python} and \proglang{Zig}, that calculates the $DFA$, $DCCA$ $\rho_{DCCA}$ and the $DMC_x^2$.
 The package presents an optimized algorithm that significantly improves the calculations speed.
-A comparison with ohter packages that calculates the .The package is also the firts to implement the $DMC_x^2$ corefficient for any number of series.
+A comparison with other packages that calculates the .The package is also the first to implement the $DMC_x^2$ coefficient for any number of series.
 }
@@
 
+% rho DCCA applications
+
 The Detrended Multiple Cross-Correlation Coefficient \citep{Zebende2018} ($DMC_x^2$) is a generalization of the $\rho_{DCCA}$ coefficient that correlates one time series (dependent variable) a number of time series (independent variables).
 The $DMC_x^2$ values ranges from zero ($0$), indicating no correlation to $1$, meaning perfect correlation or anti-correlation between the dependent and the independent variables.
+
+% DMCx2 applications
 
@@
 
+The algorithms that calculates the $\rho_{DCCA}$ uses the $DFA$ and the $DCCA$ steps.
+The $DMC_x^2$ coefficient uses the $\rho_{DCCA}$ coefficient and, consequently, also embraces the $DFA$ and the $DCCA$.
+The $DFA$ method is described in six steps:
+
+\begin{enumerate}\label{steps:dfa}
+
+  \item Taking a time series \(\{x_{i}\}\) with \(i\) ranging from \(1\) to \(N\), the integrated series \(X_{k}\) is calculated by \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) with \(k\) also ranging from \(1\) to \(N\);
+  \item the  \(X_{k}\) series is divided in \(N - n\) boxes of size \(n\)(time scale), each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
+  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting \(\widetilde{X}_{k, i}\) with \( i \le k \le (i + n) \) (detrended values);
+  \item in each box is calculated: \(f_{DFA}^{2}(n, i) = \frac{1}{1+n} \sum_{k=i}^{i + n}(X_{k}-\widetilde{X}_{k, i})^{2}\)
+  \item for all the boxes of a time scale, the $DFA$ is calculated as:\\[10pt]
+        \(F_{DFA}(n) = \sqrt{\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DFA}^{2}(n, i)}\);
+  \item for a number of different timescales (n), with possible values \( 4 \le n \le \frac{N}{4}\) the \(F_{DFA}\) function is calculated to find a relation among \(F_{DFA} \times n\)
+
+\end{enumerate}
+
+The $DCCA$ method is very similar to the $DFA$ calculations. With the difference of analyzing two series while the $DFA$ evaluate properties of a single time series. It´s also a six steps process:
+
+\begin{enumerate}
+  \label{steps:DCCA}
+  \item Taking two time series with the same length $\{x\alpha_{i}\}$ and \(\{x\beta_{i}\}\) with \(i\) ranging from \(1\) to \(N\),
+        the integrated series \(X\alpha_{k}\) and \(X\beta_{k}\) are calculated by
+        \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) for each series, with \(k\) also ranging from \(1\) to \(N\);
+  \item \(X\alpha_{k}\) and \(X\beta_{k}\) series are divided in \(N - n\) boxes of size \(n\)(time scale), each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
+  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting
+        $\widetilde{X\alpha}_{k, i}$ and $\widetilde{X\beta}_{k, i}$,
+        for series $\{x\alpha_{i}\}$ and $\{x\beta_{i}\}$ respectively,
+        with \( i \le k \le (i + n) \) (detrended values);
+  \item in each box is calculated: $f_{DCCA}^{2}(n, i) =
+        \frac{1}{1+n} \sum_{k=i}^{i + n}(X\alpha_{k}-\widetilde{X\alpha}_{k, i}) \times (X\beta_{k}-\widetilde{X\beta}_{k, i})$
+  \item for all the boxes of a time scale, the $DCCA$ is calculated as:\\[10pt]
+        $F_{DCCA}^{2}(n) =\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DCCA}^{2}(n, i)$;
+  \item for a number of different timescales (n), with possible values
+        $4 \le n \le \frac{N}{4}$ the $F_{DCCA}^{2}$ function is calculated to find a relation among
+        $F_{DCCA}^{2} \times n$
+\end{enumerate}
+
+Comparing the algorithms, the first three are basically identical, the only difference is that the $DCCA$ method apply those steps to two series.
+The step four of the $DFA$ is essentially an application of the variance calculation and the equivalent step of the $DCCA$ is a covariance between the two series.
+Step five calculates the square root of the mean of the variances calculated in  in each box for the $DFA$, in the $DCCA$, the mean of the covariances calculated for each box is calculated in stead.
+The last step,in both cases, is more a reminder to repeat the respective previous operations for a number of difference time scales.
+
+The $\rho_{DCCA}$ is measured using Eq.
+\ref{eq:p_dcca}.
+Considering the relation between $DFA$ and variance and $DCCA$ and covariance, the $\rho_{DCCA}$ resembles Pearson correlation for a time scale $n$.
+
+\begin{equation}
+  {\rho}_{DCCA}(n) = \frac{F_{DCCA~(x\alpha,~x\beta)}^{2}(n)}
+  {F_{DFA~(x\alpha)}(n) \times F_{DFA~(x\beta)}(n)}
+  \label{eq:p_dcca}
+\end{equation}
+
+The $DMC_x^2$ is a generalization of the $\rho_{DCCA}$ that calculates the correlation between one time-series $\left\lbrace Y \right\rbrace $, as the  dependent variable, and a number $j$ of time-series $\left\lbrace X_{1} \right\rbrace $, $\left\lbrace X_{2} \right\rbrace $, $\left\lbrace X_{3} \right\rbrace $, $\dots $, $\left\lbrace X_{j} \right\rbrace $ defined as independent variables.
+The coefficient is expressed mathematically as:
+
 \begin{equation}
@@
 \end{equation}
+
+In Eq.
+\ref{eq:dmcx2}, $\rho^{-1}(n)$ represent the inverse of a matrix populated by all possible combinations of $\rho_{DCCA}$ between independent variables.
+In Eq.
+\ref{eq:p_dcca_matrix}, value $\rho_{X_{1},X_{2}}(n)$, for instance, is the $\rho_{DCCA}$ for independent variables $X_{1}$ and $X_{2}$ calculated with time scale $n$, occupying position $\rho_{1 2}$ of the matrix.
+Two fundamental characteristics: the first is that the main diagonal values are all ones, since it's position in the matrix denotes the calculation of a correlation between a series and itself.
+Second, the matrix is symmetric in relation to the main diagonal, as the $\rho_{DCCA}$ is evaluated a commutative expression.
 
@@
 
-\begin{equation}
-  \rho_{Y,X_k}(n)^T=[\rho_{Y,X_1}(n), \rho_{Y,X_2}(n),\cdots,\rho_{Y,X_j}(n)]
-  \label{eq:rhocol}
+At last Eq.
+\ref{eq:rhocol} represent the transposed vector of the $\rho_{Y,X_{i}}(n)$ between the depended variable $\left\lbrace Y \right\rbrace $ and each $\left\lbrace X_{i} \right\rbrace $ independent variable for a given time scale $n$.
+
+\begin{equation} \label{eq:rhocol}
+  \rho_{Y,X_i}(n)^T=[\rho_{Y,X_1}(n), \rho_{Y,X_2}(n),\cdots,\rho_{Y,X_j}(n)]
 \end{equation}
 
-\begin{equation}
-  {\rho}_{DCCA}(n) = \frac{F_{DCCA~(x\alpha,~x\beta)}^{2}(n)}
-  {F_{DFA~(x\alpha)}(n) \times F_{DFA~(x\beta)}(n)}
-  \label{eq:p_dcca}
-\end{equation}
-
-\begin{description}
-
-  \label{steps:dfa}
-  \item Taking a time series \(\{x_{i}\}\) with \(i\) ranging from \(1\) to \(N\), the integrated series \(X_{k}\) is calculated by \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) with \(k\) also ranging from \(1\) to \(N\);
-  \item the  \(X_{k}\) series is divided in \(N - n\) boxes of size \(n\)(time scale),each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
-  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting \(\widetilde{X}_{k, i}\) with \( i \le k \le (i + n) \) (detrended values);
-  \item in each box is calculated: \(f_{DFA}^{2}(n, i) = \frac{1}{1+n} \sum_{k=i}^{i + n}(X_{k}-\widetilde{X}_{k, i})^{2}\)
-  \item for all the boxes of a time scale, the DFA is calculated as:\\[10pt]
-        \(F_{DFA}(n) = \sqrt{\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DFA}^{2}(n, i)}\);
-  \item for a number of different timescales (n), with possible values \( 4 \le n \le \frac{N}{4}\) the \(F_{DFA}\) function is calculated to find a relation among \(F_{DFA} \times n\)
-
-\end{description}
-
-\begin{description}
-  \label{steps:DCCA}
-  \item Taking two time series with the same length \(\{x\alpha_{i}\}\) and \(\{x\beta_{i}\}\) with \(i\) ranging from \(1\) to \(N\),
-        the integrated series \(X\alpha_{k}\) and \(X\beta_{k}\) are calculated by
-        \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) for each series, with \(k\) also ranging from \(1\) to \(N\);
-  \item \(X\alpha_{k}\) and \(X\beta_{k}\) series are divided in \(N - n\) boxes of size \(n\)(time scale),
-        each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
-  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting
-        \(\widetilde{X\alpha}_{k, i}\) and \(\widetilde{X\beta}_{k, i}\),
-        for series \(\{x\alpha_{i}\}\) and \(\{x\beta_{i}\}\) respectively,
-        with \( i \le k \le (i + n) \) (detrended values);
-  \item in each box is calculated: \(f_{DCCA}^{2}(n, i) =
-        \frac{1}{1+n} \sum_{k=i}^{i + n}(X\alpha_{k}-\widetilde{X\alpha}_{k, i}) \times (X\beta_{k}-\widetilde{X\beta}_{k, i})\)
-  \item for all the boxes of a time scale, the DFA is calculated as:\\[10pt]
-        \(F_{DCCA}^{2}(n) =\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DCCA}^{2}(n, i)\);
-  \item for a number of different timescales (n), with possible values
-        \( 4 \le n \le \frac{N}{4}\) the \(F_{DCCA}^{2}\) function is calculated to find a relation among
-        \(F_{DCCA}^{2} \times n\)
-\end{description}
+The $\rho_{DCCA}$ and the $DMC_x^2$ should be evaluated in a number of time scales ($n$) to analyze the characteristics of each coefficient.
 
@@
 \section{Summary and discussion} \label{sec:summary}
-
-\begin{leftbar}
-  As usual \dots
-\end{leftbar}
 

### 2024-11-06 paper_v01/article.tex
update flowchart

@@
 Section \ref{sec:optimization} shows how this library was implemented, the optimization technics and the recommended steps to use the library.
-In Section \ref{sec:results} the \pkg{Zebende} package is compared with other packages for \proglang{Python} and \proglang{R} that calculates the $\rho_{DCCA}$ in terms of performance and usability, leading to the conclusions in \ref{sec:summary}.
+In Section \ref{sec:results} the \pkg{Zebende} package is compared with other packages for \proglang{Python} and \proglang{R} that calculates the $\rho_{DCCA}$ in terms of performance and usability, leading to the conclusions in Section \ref{sec:summary}.
 
@@
 \section{Zebende package: implementation and optimization} \label{sec:optimization}
+
+The implementation of the \pkg{Zebende} package follows some well defined golas:
+
+\begin{enumerate}
+  \item Enhance performance;
+  \item avoid redundant calculations;
+  \item make the outputs compatibles with other data analyses tools (including data manipulation, machine learn and statistical packages);
+  \item deliver a customizable set of tools;
+  \item facilitate package evolution and maintenance;
+  \item create an easy to use package.
+\end{enumerate}
+
+The \proglang{Python} language was chosen because it´s one of the most used languages in the data analyses field and have a great support for statistical tools and machine learning algorithms.
+The are a pletora of tools to load and manipulate data (\pkg{Pandas}, \pkg{Polar}, \pkg{PySpark} ...), execute statistical analyzes (\pkg{Numpy}, \pkg{SciPy}, \pkg{StatsModels} ...), machine learning (\pkg{Pytorch}, \pkg{TensorFlow}, \pkg{Scykit Learn}...) and data visualization(\pkg(Matplotlib), \pkg{Seaborn}, \pkg{Plotly}) among other related applications.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics{./figs/pdcca_chart.png}
+  \caption{\label{fig:pdcca_chart}Calculating  $\rho_{DCCA}$ with \pkg{Zebende} package - Simplified flowchart}
+\end{figure}
+
+The first draft of the code was written in pure \proglang{Python}, to rapidly prototype the way users will interact with the package.
+Figures \ref{fig:pdcca_chart} and \ref{fig:dmc_chart} presents simplified flowcharts illustrating how to use the package and how the main functions ($\rho_{DCCA}$ and $DMA_x^2$) works.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics{./figs/dmc_chart.png}
+  \caption{\label{fig:dmc_chart}Calculating  $DMC_x^2$ with \pkg{Zebende} package - Simplified flowchart}
+\end{figure}
+
+The preparation steps are the same in both functions.
+First the data is loaded, and should be analyzed by the researchers.
+Base on the data characteristics, the set should be treated to ensure the methods requirements in the "Pre-processing" stage.
+The package functions expects data as ma matrix with the columns as the series and the lines as time steps.
+Columns unused in the indented research should also be dropped for better performance os the algorithms in this step.
+The more common way to do that is to use a data manipulation package.
+To proceed to the next step, the data table should be in the form of a \pkg{Numpy} 2D array, and this could be done with any data manipulation \proglang{Python} package.
+The next step is to calculate the integrated series.
+The package provides a function, named \verb"integrated_series()", to calculate that.
+The code example below show how to load the libraries (using \pkg{Pandas} as the data manipulation packages and loading a \verb".csv" file as a generic example), convert to \pkg{Numpy} array and generate the integrated series.
+
+\begin{Code}
+# importing packages
+import numpy as np
+import pandas as pd
+import zebende as zb
+
+data = pd.read_csv('path_to_the_file.csv') # loading data
+# Pre-processing data
+# ...
+data = data.to_numpy(data) #converting data to Numpy array
+int_data = zb.integrated_series(data) # calculating the integrated series
+\end{Code}
+
+The option of taking out the integrated series generation from the main functions to an independent one was inspired by \cite{Peng_1994} work, where the way of calculating integrated series was different from the one that is widely used in more recent years.
+The integration of the series is essentially a pre-processing step, and this approach makes easy to explore alternative ways to integrate the series or compare \cite{Peng_1994} approach to the current one in different scenarios.
+
+The input and output structures of each function are displayed below:
+
+\begin{Code}
+def p_dcca(
+    input_data: NDArray[float64],
+    tws: NDArray[int64] | NDArray[float64],
+    DCCA_of: ndarray | Literal['all'] = "all",
+    P_DCCA_output_matrix: bool = False
+) -> tuple[NDArray[float64],  # DFA
+          NDArray[float64],   # DCCA
+          NDArray[float64]    # P_DCCA
+          ]
+\end{Code}
+
+\begin{Code}
+def dmcx2(
+    input_data: NDArray[float64],
+    tws: NDArray[int64] | NDArray[float64],
+    dmcx2_of: ENUM_DMCx2_of | NDArray[float64] | list = 'all-full'
+) -> tuple[ NDArray[float64],   # DFA
+            NDArray[float64],   # DCCA
+            NDArray[float64],   # P_DCCA
+            NDArray[float64],   # DMC
+          ]
+  \end{Code}
+
+The first two inputs are the same for functions \verb"p_dcca()" and \verb"dmcx2()": \verb"input_data" receive the integrated series and the \verb"tws" receives an 1D array representing the time scales (box size) described in the algorithms on Section \ref{sec:calculations}.
+The \verb"input_data" is a 2D array of 64 bits floating point data.
+The \verb"tws" accepts integers and, for convenience reasons, also floating points.
+Since the size of the boxes needs to be integers, in case of floating points, the values will be converted to integers by ignoring the decimal values(truncating).
+This two inputs are colored in light gray in Figures \ref{fig:pdcca_chart} and \ref{eq:dmcx2}, indicating mandatory inputs.
+
+With the mandatory steps explained, some very important optional inputs should be addressed.
+Starting with the $\rho_{DCCA}$ function (dark green node in Figure \ref{fig:dmc_chart}) represents the input \verb"DCCA_of" of the function.
+This input requires a 2D array of index.
+For example: if the \verb"input_data" receives a four columns matrix, with index ranging from $0$ to $3$, that is intended to calculate de $DFA$ for all the series but the $\rho_{DCCA}$ between series of index $0$ and $1$ and also for series $2$ and $3$, the \verb"DCCA_of" input should receive the array \verb"[[0,1], [2,3]]".
+If no value (or the string \verb"'all'") is given, the function will calculate all possible combinations of $DCCA$ calculations between all the series respecting the index values order, as below:
+
+\begin{Code}
+[[0,1], [0,2], [0,3], [1,2], [1,3], [2,3]]
+\end{Code}
+
+It´s important to understand the calculation steps, the role of the \verb"DCCA_of" array and how it fits in the goals of the package implementation.
+The code below is part of the \proglang{Python} implementation of the $\rho_{DCCA}$ function,
+
+\begin{Code}
+for n_index, n  in enumerate(tws): # for each time scale
+  # temporaty allocation arrays
+  f2dfa_n = np.full(shape=(data.shape[0] - n, data.shape[1]),
+  fill_value=np.nan, dtype=data.dtype)
+  dcca_n = np.full(shape=(data.shape[0] - n, DCCA_of.shape[0]), 
+  fill_value=np.nan, dtype=data.dtype)
+  detrended_mat = np.full(shape=(n + 1, data.shape[1]), 
+  fill_value=np.nan, dtype=data.dtype)
+  for i in range(data.shape[0] - n): # for each box
+      detrended_series( # inputs
+                        time_steps[i : i + (n + 1)],  # arr_x
+                        data[i : i + (n + 1), :],  # mat_y
+                         detrended_mat,  # output
+                      )
+      f2dfa_n[i] = np.power(detrended_mat, 2).mean(axis=0)
+      for j, pair in enumerate(DCCA_of): # for each DCCA pair
+          dcca_n[i, j] = (detrended_mat[:, pair[0]] * detrended_mat[:, pair[1]]
+                         ).mean(axis=0)
+  F_DFA_arr[n_index, :] = np.sqrt(f2dfa_n.mean(axis=0))
+  DCCA_arr[n_index, :] = dcca_n.mean(axis=0)
+  # calculation of P_DCCA
+  P_DCCA_output_function(n_index, DCCA_of, F_DFA_arr, DCCA_arr, P_DCCA_arr)
+\end{Code}
+
+The first \verb"for" loop in the code operates over the values of the \verb"tws" input, asserting that the step 6 of the $DFA$ and $DCCA$ methods in Section \ref{sec:calculations}.
+Three temporary arrays are allocated for each time scale, \verb"f2dfa_n"
 

### 2024-11-06 paper_v01/article.tex
P_DCCA function explained

@@
 Starting with the $\rho_{DCCA}$ function (dark green node in Figure \ref{fig:dmc_chart}) represents the input \verb"DCCA_of" of the function.
-This input requires a 2D array of index.
+This input requires a 2D array of integers, each row is a pair of index, related to the \verb"data_input" matrix.
 For example: if the \verb"input_data" receives a four columns matrix, with index ranging from $0$ to $3$, that is intended to calculate de $DFA$ for all the series but the $\rho_{DCCA}$ between series of index $0$ and $1$ and also for series $2$ and $3$, the \verb"DCCA_of" input should receive the array \verb"[[0,1], [2,3]]".
@@
   # calculation of P_DCCA
-  P_DCCA_output_function(n_index, DCCA_of, F_DFA_arr, DCCA_arr, P_DCCA_arr)
+  P_DCCA_output_function(n_index, DCCA_of, F_DFA_arr, DCCA_arr, # Inputs
+  P_DCCA_arr) # Output
 \end{Code}
@@
 The first \verb"for" loop in the code operates over the values of the \verb"tws" input, asserting that the step 6 of the $DFA$ and $DCCA$ methods in Section \ref{sec:calculations}.
-Three temporary arrays are allocated for each time scale, \verb"f2dfa_n"
+Three temporary arrays are allocated for each time scale.
+The first, \verb"f2dfa_n", to store the calculations of the step 4 of the $DFA$.
+The number of lines of this array correspond to the number of boxes in the current time scale ($N - n$) and the columns is the number of series in the analysis.
+Array  \verb"dcca_n", holds the values calculated in the step 4 of the $DCCA$.
+The number of lines also correspond to the number of boxes but the number of columns equals the number of pairs (rows) in the \verb"DCCA_of" input.
+The \verb"detrended_mat" array has a number of lines is the number of points in a time scale box ($n +1$) and column count also equal to the number of time series.
+This last array is very important to the implementation goals.
+
+In each box, the \verb"detrended_series" function execute a one degree polynomial fit, subtract the values of the serie with the value given by the interpolated curve and stores this values in the \verb"detrended_mat".
+After the $DFA$ is calculated.
+With all the detrended values calculated, another loop, for all the rows in the \verb"DCCA_of", uses the \verb"detrended_mat" and the $DFA$ arrays to calculate the $DCCA$.
+This approach avoid redundant calculation.
+If the \verb"DCCA_of" is $[[0,1], [0,2], [1,2]]$, the detrended and $DFA$ values for each serie are calculated just once.
+If in a set of four series, a research intend to calculate the $DCCA$ between the first serie and all the others along with the two last series, inputting  \verb"DCCA_of" $= [[0,1], [0,2], [0,3], [2,3]]$.
+In this case, the \verb"detrended_mat" will be calculated just once for each series and the $DCCA$ between series $[[1,2], [1,3]]$ will be skipped, revealing a very customizable algorithm.
+
+The $\rho_{DCCA}$ function, named \verb"P_DCCA_output_function" in the code is, in fact a pointer to other functions.
+One that outputs the  $\rho_{DCCA}$ results in the form of a table (rows for each time scale and columns for each \verb"DCCA_of" pair) and the other outputs it the form of a 3D matrix, where each level is the matrix in Equation \ref{eq:p_dcca_matrix} for one of the time scales.
+This behavior is driven by the \verb"P_DCCA_output_matrix" input (represented as the light green node in Figure \ref{fig:pdcca_chart}), where \verb"False" means table output and \verb"True" matrix output.
+This is very convenient for calculating the $DMC_x^2$.
+
+The $DMC_x^2$ function runs the $\rho_{DCCA}$ with \verb"P_DCCA_output_matrix" set to \verb"True".
 

### 2024-11-07 paper_v01/article.tex
P_DCCA function explained

@@
 The $DMC_x^2$ function runs the $\rho_{DCCA}$ with \verb"P_DCCA_output_matrix" set to \verb"True".
+There is no \verb"DCCA_of" input for the $DMC_x^2$ function, instead there is a \verb"dmcx2_of" parameter.
+This input receives a 2D matrix where each line represents the indexes of the series to be used in Equation \ref{eq:dmcx2}.
+The first value of each row is the index of the serie used as the dependent variable, the others, the independent ones.
+There are two literal strings that can be used, for convenience as inputs for this parameter: \verb"'all-full'", that generates a 2D array with every series as the dependent variable against all the others; and \verb"'first-full'", with only one row, having the index zero series as the dependente variable in relation with all the others.
 

### 2024-11-15 paper_v01/article.tex
P_DCCA function explained

@@
 
-  \item Taking a time series \(\{x_{i}\}\) with \(i\) ranging from \(1\) to \(N\), the integrated series \(X_{k}\) is calculated by \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) with \(k\) also ranging from \(1\) to \(N\);
-  \item the  \(X_{k}\) series is divided in \(N - n\) boxes of size \(n\)(time scale), each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
-  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting \(\widetilde{X}_{k, i}\) with \( i \le k \le (i + n) \) (detrended values);
-  \item in each box is calculated: \(f_{DFA}^{2}(n, i) = \frac{1}{1+n} \sum_{k=i}^{i + n}(X_{k}-\widetilde{X}_{k, i})^{2}\)
+  \item Taking a time series $\{x_{i}\}$ with $i$ ranging from $1$ to $N$, the integrated series $X_{k}$ is calculated by $X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right]$ with \(k\) also ranging from $1$ to $N$;
+  \item the  $X_{k}$ series is divided in $N - n$ boxes of size $n$(time scale), each box containing $n + 1$ values, starting in $i$ up to $i + n$;
+  \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting $\widetilde{X}_{k, i}$ with $i \le k \le (i + n)$;
+  \item in each box is calculated: $f_{DFA}^{2}(n, i) = \frac{1}{1+n} \sum_{k=i}^{i + n}(X_{k}-\widetilde{X}_{k, i})^{2}$
   \item for all the boxes of a time scale, the $DFA$ is calculated as:\\[10pt]
-        \(F_{DFA}(n) = \sqrt{\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DFA}^{2}(n, i)}\);
-  \item for a number of different timescales (n), with possible values \( 4 \le n \le \frac{N}{4}\) the \(F_{DFA}\) function is calculated to find a relation among \(F_{DFA} \times n\)
+        $F_{DFA}(n) = \sqrt{\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DFA}^{2}(n, i)}$;
+  \item for a number of different timescales ($n$), with possible values $4 \le n \le \frac{N}{4}$ the $F_{DFA}$ function is calculated to find a relation among $F_{DFA} \times n$
 
@@
 
-The $DCCA$ method is very similar to the $DFA$ calculations. With the difference of analyzing two series while the $DFA$ evaluate properties of a single time series. It´s also a six steps process:
+The $DCCA$ method is very similar to the $DFA$ calculations, with the difference of analyzing two series while the $DFA$ evaluate properties of a single time series. It´s also a six steps process:
 
@@
   \label{steps:DCCA}
-  \item Taking two time series with the same length $\{x\alpha_{i}\}$ and \(\{x\beta_{i}\}\) with \(i\) ranging from \(1\) to \(N\),
-        the integrated series \(X\alpha_{k}\) and \(X\beta_{k}\) are calculated by
-        \(X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] \) for each series, with \(k\) also ranging from \(1\) to \(N\);
-  \item \(X\alpha_{k}\) and \(X\beta_{k}\) series are divided in \(N - n\) boxes of size \(n\)(time scale), each box containing \(n + 1\) values, starting in \(i\) up to \(i + n\);
+  \item Taking two time series with the same length $\{x\alpha_{i}\}$ and $\{x\beta_{i}\}$ with $i$ ranging from $1$ to $N$,
+        the integrated series $X\alpha_{k}$ and $X\beta_{k}$ are calculated by
+        $X_{k} = \sum_{i=1}^{k}\left[x_{i} - \langle x \rangle \right] $ for each series, with $k$ also ranging from $i$ to $N$;
+  \item $X\alpha_{k}$ and $X\beta_{k}$ series are divided in $N - n$ boxes of size $n$(time scale), each box containing $n + 1$ values, starting in $i$ up to $i + n$;
   \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting
@@
         for series $\{x\alpha_{i}\}$ and $\{x\beta_{i}\}$ respectively,
-        with \( i \le k \le (i + n) \) (detrended values);
+        with $i \le k \le (i + n)$;
   \item in each box is calculated: $f_{DCCA}^{2}(n, i) =
@@
         $F_{DCCA}^{2}(n) =\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DCCA}^{2}(n, i)$;
-  \item for a number of different timescales (n), with possible values
+  \item for a number of different timescales ($n$), with possible values
         $4 \le n \le \frac{N}{4}$ the $F_{DCCA}^{2}$ function is calculated to find a relation among
@@
 Comparing the algorithms, the first three are basically identical, the only difference is that the $DCCA$ method apply those steps to two series.
-The step four of the $DFA$ is essentially an application of the variance calculation and the equivalent step of the $DCCA$ is a covariance between the two series.
-Step five calculates the square root of the mean of the variances calculated in  in each box for the $DFA$, in the $DCCA$, the mean of the covariances calculated for each box is calculated in stead.
+The step four of the $DFA$ can be considered an analogous of the variance, replacing the average subtraction (in the variance) for the values obtained by the polinomial fit(estimated seires); and the equivalent step of the $DCCA$ is, in teh same terms, compared to the covariance between the two series.
+Step five calculates the square root of the mean of the values calculates in the previous step $(f_{DFA}^{2}$ in each box for the $DFA$, in the $DCCA$, the mean of the values evaluated  each box is calculated in stead.
 The last step,in both cases, is more a reminder to repeat the respective previous operations for a number of difference time scales.
@@
   \item make the outputs compatibles with other data analyses tools (including data manipulation, machine learn and statistical packages);
-  \item deliver a customizable set of tools;
+  \item create a customizable and modular set of tools;
   \item facilitate package evolution and maintenance;
-  \item create an easy to use package.
+  \item deliver an easy to use package.
 \end{enumerate}
@@
 
-The $\rho_{DCCA}$ function, named \verb"P_DCCA_output_function" in the code is, in fact a pointer to other functions.
+The function named \verb"P_DCCA_output_function" in the code above,is a pointer to other functions.
 One that outputs the  $\rho_{DCCA}$ results in the form of a table (rows for each time scale and columns for each \verb"DCCA_of" pair) and the other outputs it the form of a 3D matrix, where each level is the matrix in Equation \ref{eq:p_dcca_matrix} for one of the time scales.
@@
 This is very convenient for calculating the $DMC_x^2$.
+There are two utility functions to transform a table output in a matrix one (\verb"p_dcca_simple_to_matrix") and also the other way around (\verb"p_dcca_matrix_to_simple").
 
-The $DMC_x^2$ function runs the $\rho_{DCCA}$ with \verb"P_DCCA_output_matrix" set to \verb"True".
+The \verb"dmcx2" function runs the \verb"p_dcca()" with \verb"P_DCCA_output_matrix" set to \verb"True".
 There is no \verb"DCCA_of" input for the $DMC_x^2$ function, instead there is a \verb"dmcx2_of" parameter.
@@
 There are two literal strings that can be used, for convenience as inputs for this parameter: \verb"'all-full'", that generates a 2D array with every series as the dependent variable against all the others; and \verb"'first-full'", with only one row, having the index zero series as the dependente variable in relation with all the others.
+The \verb"'all-full'" option is operated by the function \verb"dmc_of_all_as_y".
+
+There is a restriction to the \verb"dmcx2_of" array: the independent values must be in crescent order.
+There is a check in the \verb"dmcx2" that raises an error if the values are out of order, and suggests the use of a utility function named \verb"ordering_x_dmcx2_of".
+The code below shows an example of a manually defined \verb"dmcx2_of" array with 2 elements out of order in the second row, the \verb"AssertionError" output, the way of fixing it with the \verb"ordering_x_dmcx2_of" function and finally the outputted ordered array.
+
+\begin{CodeChunk}
+  \begin{CodeInput}
+dmcx2_of = np.array([ [0, 1, 2, 3],
+                      [1, 0, 3, 2],
+                      [2, 0, 1, 3],
+                      [3, 0, 1, 2]  ])
+
+dfa , dcca, pdcca, dmc = zb.dmcx2(int_data, tws, dmcx2_of=dmcx2_of)
+  \end{CodeInput}
+  \begin{CodeOutput}
+AssertionError: 
+dmcx2_of x values out of order: 
+use zebende.ordering_x_dmcx2_of(dmcx2_of)
+ to fix it before passing the dmcx2_of values to zebende.dmcx2() function"
+  \end{CodeOutput}
+  \begin{CodeInput}
+dmcx2_of = ordering_x_dmcx2_of(dmcx2_of)
+print(dmcx2_of)
+  \end{CodeInput}
+  \begin{CodeOutput}
+[[0 1 2 3]
+ [1 0 2 3]
+ [2 0 1 3]
+ [3 0 1 2]]
+  \end{CodeOutput}
+\end{CodeChunk}
 

### 2024-11-16 paper_v01/article.tex
corrections and DV concept

@@
 
-The $\rho_{DCCA}$~\citep{Zebende2011} is a widely used coefficient that measures the cross-correlation between tow non-stationary time series.
+The Detrended Cross-correlation Coefficient ($\rho_{DCCA}$)~\citep{Zebende2011} is a widely used coefficient that measures the cross-correlation between tow non-stationary time series.
 It´s an extension of the Detrended Fluctuation Analysis ($DFA$)~\citep{Peng_1994} and the Detrended Cross-correlation Analysis ($DCCA$)~\citep{Podobnik2008}: while the $DFA$ calculates the self-affinity and long-memory properties of a time series data, and the $DCCA$ analyses power-law cross correlations between two different non-stationarity time series, the $\rho_{DCCA}$ coefficient quantifies this cross-correlation in simple values ranging from $-1$ to $1$, where $-1$ indicates a perfect anti-correlation between the series, $1$ a perfect correlation and zero ($0$) no correlation at all.
@@
   \item for each box, a polynomial (usually of degree 1) best fit is calculated, getting $\widetilde{X}_{k, i}$ with $i \le k \le (i + n)$;
-  \item in each box is calculated: $f_{DFA}^{2}(n, i) = \frac{1}{1+n} \sum_{k=i}^{i + n}(X_{k}-\widetilde{X}_{k, i})^{2}$
+  \item in each box is calculated: $f_{DFA}^{2}(n, i) = \frac{1}{1+n} \sum_{k=i}^{i + n}(X_{k}-\widetilde{X}_{k})^{2}$
   \item for all the boxes of a time scale, the $DFA$ is calculated as:\\[10pt]
@@
   \item in each box is calculated: $f_{DCCA}^{2}(n, i) =
-          \frac{1}{1+n} \sum_{k=i}^{i + n}(X\alpha_{k}-\widetilde{X\alpha}_{k, i}) \times (X\beta_{k}-\widetilde{X\beta}_{k, i})$
+          \frac{1}{1+n} \sum_{k=i}^{i + n}(X\alpha_{k}-\widetilde{X\alpha}_{k, i}) \times (X\beta_{k}-\widetilde{X\beta}_{k})$
   \item for all the boxes of a time scale, the $DCCA$ is calculated as:\\[10pt]
@@
 Comparing the algorithms, the first three are basically identical, the only difference is that the $DCCA$ method apply those steps to two series.
-The step four of the $DFA$ can be considered an analogous of the variance, replacing the average subtraction (in the variance) for the values obtained by the polinomial fit(estimated seires); and the equivalent step of the $DCCA$ is, in teh same terms, compared to the covariance between the two series.
-Step five calculates the square root of the mean of the values calculates in the previous step $(f_{DFA}^{2}$ in each box for the $DFA$, in the $DCCA$, the mean of the values evaluated  each box is calculated in stead.
-The last step,in both cases, is more a reminder to repeat the respective previous operations for a number of difference time scales.
+The step four of the $DFA$ can be considered an analogous of the variance, replacing the average subtraction (in the variance) for the values obtained by the polinomial fit(estimated seires); and the equivalent step of the $DCCA$ is, in the same terms, compared to the covariance between the two series.
+The technique of fitting a curve, interpreted as a trend inside each box, and subtracting the value estimated by the trend from the actual value in the integrated series $(X_{k}-\widetilde{X}_{k})$ from now on will be called \textbf{detrended values}($DV$).
+In the $DFA$ algorithm, the $f_{DFA}^{2}(n, i)$ function is the mean of the square of the $DV$, in $DCCA$ calculations, the $f_{DCCA}^{2}(n, i)$ function evaluates the mean of the product of the $DV$ of the two series in each box.
+
+Step five of the $DFA$ calculates the square root of the mean of the values calculates in the previous step for each box, in the $DCCA$, the mean of the values evaluated for each box is calculated in stead.
+The last step, in both cases, is more a reminder to repeat the respective previous operations for a number of difference time scales ($n$).
 
@@
 \ref{eq:p_dcca_matrix}, value $\rho_{X_{1},X_{2}}(n)$, for instance, is the $\rho_{DCCA}$ for independent variables $X_{1}$ and $X_{2}$ calculated with time scale $n$, occupying position $\rho_{1 2}$ of the matrix.
-Two fundamental characteristics: the first is that the main diagonal values are all ones, since it's position in the matrix denotes the calculation of a correlation between a series and itself.
-Second, the matrix is symmetric in relation to the main diagonal, as the $\rho_{DCCA}$ is evaluated a commutative expression.
+Two fundamental characteristics: the first is that the main diagonal values are all ones, since it's position in the matrix denotes the calculation of a cross-correlation between a series and itself.
+Second, the matrix is symmetric in relation to the main diagonal, as the $\rho_{DCCA}$ is a commutative operation.
 
@@
 
-The $\rho_{DCCA}$ and the $DMC_x^2$ should be evaluated in a number of time scales ($n$) to analyze the characteristics of each coefficient.
+As the $DFA$ and the $DCCA$,~$\rho_{DCCA}$ and~$DMC_x^2$ should be evaluated in a number of time scales ($n$) to analyze the characteristics of each coefficient.
 
@@
 The \proglang{Python} language was chosen because it´s one of the most used languages in the data analyses field and have a great support for statistical tools and machine learning algorithms.
-The are a pletora of tools to load and manipulate data (\pkg{Pandas}, \pkg{Polar}, \pkg{PySpark} ...), execute statistical analyzes (\pkg{Numpy}, \pkg{SciPy}, \pkg{StatsModels} ...), machine learning (\pkg{Pytorch}, \pkg{TensorFlow}, \pkg{Scykit Learn}...) and data visualization(\pkg(Matplotlib), \pkg{Seaborn}, \pkg{Plotly}) among other related applications.
+The are a pletora of tools to load and manipulate data (\pkg{Pandas}, \pkg{Polar}, \pkg{PySpark} ...), execute statistical analyzes (\pkg{Numpy}, \pkg{SciPy}, \pkg{StatsModels} ...), machine learning (\pkg{Pytorch}, \pkg{TensorFlow}, \pkg{Scykit Learn}...) and data visualization(\pkg(Matplotlib), \pkg{Seaborn}, \pkg{Plotly} ...) among other related applications.
 

### 2025-01-29 paper_v01/article.tex
paper version 02

@@
 The \proglang{Python} language was chosen because it´s one of the most used languages in the data analyses field and have a great support for statistical tools and machine learning algorithms.
-The are a pletora of tools to load and manipulate data (\pkg{Pandas}, \pkg{Polar}, \pkg{PySpark} ...), execute statistical analyzes (\pkg{Numpy}, \pkg{SciPy}, \pkg{StatsModels} ...), machine learning (\pkg{Pytorch}, \pkg{TensorFlow}, \pkg{Scykit Learn}...) and data visualization(\pkg(Matplotlib), \pkg{Seaborn}, \pkg{Plotly} ...) among other related applications.
+The are a pletora of tools to load and manipulate data (\pkg{Pandas}, \pkg{Polar}, \pkg{PySpark} ...), execute statistical analyzes (\pkg{Numpy}, \pkg{SciPy}, \pkg{StatsModels}~...), machine learning (\pkg{Pytorch}, \pkg{TensorFlow}, \pkg{Scykit Learn}~...) and data visualization(\pkg{Matplotlib}, \pkg{Seaborn}, \pkg{Plotly}~...) among other data related applications.
 
@@
 First the data is loaded, and should be analyzed by the researchers.
-Base on the data characteristics, the set should be treated to ensure the methods requirements in the "Pre-processing" stage.
-The package functions expects data as ma matrix with the columns as the series and the lines as time steps.
-Columns unused in the indented research should also be dropped for better performance os the algorithms in this step.
+Based on the data characteristics, the set should be treated to ensure the methods requirements in the "Pre-processing" stage.
+The package functions expects data as a matrix with the columns as the series and the lines as time steps.
+Columns unwanted in the indented research should also be dropped for better performance of the algorithms in this step.
 The more common way to do that is to use a data manipulation package.
-To proceed to the next step, the data table should be in the form of a \pkg{Numpy} 2D array, and this could be done with any data manipulation \proglang{Python} package.
+To proceed to the next step, the data table should be in the form of a \pkg{Numpy} 2D array.
+Any data manipulation \proglang{Python} package can export a table as a \pkg{Numpy} array.
 The next step is to calculate the integrated series.
@@
 
-The option of taking out the integrated series generation from the main functions to an independent one was inspired by \cite{Peng_1994} work, where the way of calculating integrated series was different from the one that is widely used in more recent years.
-The integration of the series is essentially a pre-processing step, and this approach makes easy to explore alternative ways to integrate the series or compare \cite{Peng_1994} approach to the current one in different scenarios.
+The option of taking out the integrated series generation from the main methods (\verb"p_dcca()" and \verb"dmcx2()") to an independent one was inspired by \cite{Peng_1994} work, where the way of calculating integrated series was different from the one that is widely used in more recent years.
+The integration of the series is essentially a pre-processing step, and this approach makes easy to explore alternative ways to integrate the series or compare \cite{Peng_1994} approach to the current most used one in different scenarios, or even embrace new proposals for the series integrating step.
 

### 2025-01-29 paper_v02/article.tex
paper version 02

--- 2025-01-29 paper_v01/article.tex
@@
 The package presents an optimized algorithm that significantly improves the calculations speed.
-A comparison with other packages that calculates the .The package is also the first to implement the $DMC_x^2$ coefficient for any number of series.
+A comparison with other packages that calculates the $\rho_{DCCA}$.
+The package is also the first to implement the $DMC_x^2$ coefficient in a package and the first algorithm to calculates it for any number of time series.
 }
@@
 
-The Detrended Cross-correlation Coefficient ($\rho_{DCCA}$)~\citep{Zebende2011} is a widely used coefficient that measures the cross-correlation between tow non-stationary time series.
+The Detrended Cross-correlation Coefficient ($\rho_{DCCA}$)~\citep{Zebende2011} is a widely used coefficient that measures the cross-correlation between two non-stationary time series.
 It´s an extension of the Detrended Fluctuation Analysis ($DFA$)~\citep{Peng_1994} and the Detrended Cross-correlation Analysis ($DCCA$)~\citep{Podobnik2008}: while the $DFA$ calculates the self-affinity and long-memory properties of a time series data, and the $DCCA$ analyses power-law cross correlations between two different non-stationarity time series, the $\rho_{DCCA}$ coefficient quantifies this cross-correlation in simple values ranging from $-1$ to $1$, where $-1$ indicates a perfect anti-correlation between the series, $1$ a perfect correlation and zero ($0$) no correlation at all.
@@
   \item make the outputs compatibles with other data analyses tools (including data manipulation, machine learn and statistical packages);
+  \item manage multiple time series inputs;
+  \item operate the $DMC_x^2$ for any number of series;
   \item create a customizable and modular set of tools;
@@
 The first draft of the code was written in pure \proglang{Python}, to rapidly prototype the way users will interact with the package.
-Figures \ref{fig:pdcca_chart} and \ref{fig:dmc_chart} presents simplified flowcharts illustrating how to use the package and how the main functions ($\rho_{DCCA}$ and $DMA_x^2$) works.
+Figures \ref{fig:pdcca_chart} and \ref{fig:dmc_chart} presents simplified flowcharts illustrating how to use the package and how the main functions ($\rho_{DCCA}$ and $DMC_x^2$) works.
 
@@
 The \verb"tws" accepts integers and, for convenience reasons, also floating points.
-Since the size of the boxes needs to be integers, in case of floating points, the values will be converted to integers by ignoring the decimal values(truncating).
+Since the size of the boxes needs to be integers, in case of floating points, the values will be converted to integers by ignoring the decimal values (truncating).
 This two inputs are colored in light gray in Figures \ref{fig:pdcca_chart} and \ref{eq:dmcx2}, indicating mandatory inputs.
@@
 It´s important to understand the calculation steps, the role of the \verb"DCCA_of" array and how it fits in the goals of the package implementation.
-The code below is part of the \proglang{Python} implementation of the $\rho_{DCCA}$ function,
+The code below is part of the pure \proglang{Python} implementation of the $\rho_{DCCA}$ function,
 
@@
 
-The first \verb"for" loop in the code operates over the values of the \verb"tws" input, asserting that the step 6 of the $DFA$ and $DCCA$ methods in Section \ref{sec:calculations}.
-Three temporary arrays are allocated for each time scale.
-The first, \verb"f2dfa_n", to store the calculations of the step 4 of the $DFA$.
-The number of lines of this array correspond to the number of boxes in the current time scale ($N - n$) and the columns is the number of series in the analysis.
-Array  \verb"dcca_n", holds the values calculated in the step 4 of the $DCCA$.
-The number of lines also correspond to the number of boxes but the number of columns equals the number of pairs (rows) in the \verb"DCCA_of" input.
-The \verb"detrended_mat" array has a number of lines is the number of points in a time scale box ($n +1$) and column count also equal to the number of time series.
-This last array is very important to the implementation goals.
+The first \verb"for" loop in the code operates over the values of the \verb"tws" input, ensuring that step 6 of the $DFA$ and $DCCA$ methods, presented in Section \ref{sec:calculations}, is being carried out.
+In other words, the calculations will be applied, sequentially, to every single value in the \verb"tws" array.
+Three temporary arrays are allocated and resized for each time scale.
+The first, \verb"f2dfa_n", is used to store the calculations of the step 4 of the $DFA$~($f_{DFA}^{2}$).
+The number of lines of this array correspond to the number of boxes in the current time scale ($N - n$, resized for each time scale $n$) and the columns is the number of series in the analysis (same size for every value of $n$).
+Second, the array  \verb"dcca_n" holds the values calculated in the step 4 of the $DCCA$~(to calculate $f_{DCCA}^{2}(n, i)$).
+The number of lines also correspond to the number of boxes(resized for each $n$) but the number of columns equals the number of pairs (rows) in the \verb"DCCA_of" input (same size in each $n$).
+The \verb"detrended_mat" array has a number of lines equal to the number of points in a time scale box ($n +1$, resized for each $n$) and column count also equal to the number of time series (same size in each $n$).
+The first two tempoaray arrays will store data for all the boxes in the time scale, the last one will be used in each box and will have the values replaced in the next one, until the last box of the time scale.
+Than all the arrays will be cleared and recreated with shapes calculated with the new value of $n$.
 
-In each box, the \verb"detrended_series" function execute a one degree polynomial fit, subtract the values of the serie with the value given by the interpolated curve and stores this values in the \verb"detrended_mat".
-After the $DFA$ is calculated.
-With all the detrended values calculated, another loop, for all the rows in the \verb"DCCA_of", uses the \verb"detrended_mat" and the $DFA$ arrays to calculate the $DCCA$.
-This approach avoid redundant calculation.
-If the \verb"DCCA_of" is $[[0,1], [0,2], [1,2]]$, the detrended and $DFA$ values for each serie are calculated just once.
-If in a set of four series, a research intend to calculate the $DCCA$ between the first serie and all the others along with the two last series, inputting  \verb"DCCA_of" $= [[0,1], [0,2], [0,3], [2,3]]$.
-In this case, the \verb"detrended_mat" will be calculated just once for each series and the $DCCA$ between series $[[1,2], [1,3]]$ will be skipped, revealing a very customizable algorithm.
+After the allocation of the temporary arrays, the second \verb"for" loop operates in every box for a certain time scale $n$.
+In each box, the \verb"detrended_series()" function execute a one degree polynomial fit, subtract the values of the serie with the value given by the interpolated curve($DV$) and stores this values in the \verb"detrended_mat".
+After that, the \textbf{mean of the square of each $DV$ in the box} ($f_{DFA}^{2}(n, i)$) is calculated in the current box for every time series and stored in the \verb"f2dfa_n" array in the line associated with the current box, and the column related to each time series.
 
-The function named \verb"P_DCCA_output_function" in the code above,is a pointer to other functions.
+The next nested \verb"for" loop operates over the \verb"DCCA_of" array.
+For each line (pair) in the \verb"DCCA_of" array, the $f_{DCCA}^{2}(n, i)$ is evaluated by multiplying the corresponding $DV$ from the two series  boxes in the current pair, getting the mean, and stores it in the \verb"dcca_n" temporary array.
+
+After all the pairs are calculated, the algorithm goes back to the
+
+The function named \verb"P_DCCA_output_function()" in the code above,is a pointer to other functions.
 One that outputs the  $\rho_{DCCA}$ results in the form of a table (rows for each time scale and columns for each \verb"DCCA_of" pair) and the other outputs it the form of a 3D matrix, where each level is the matrix in Equation \ref{eq:p_dcca_matrix} for one of the time scales.
@@
 This is very convenient for calculating the $DMC_x^2$.
-There are two utility functions to transform a table output in a matrix one (\verb"p_dcca_simple_to_matrix") and also the other way around (\verb"p_dcca_matrix_to_simple").
+There are two utility functions to transform a table output in a matrix one (\verb"p_dcca_simple_to_matrix()") and also the other way around (\verb"p_dcca_matrix_to_simple()").
 
@@
 There are two literal strings that can be used, for convenience as inputs for this parameter: \verb"'all-full'", that generates a 2D array with every series as the dependent variable against all the others; and \verb"'first-full'", with only one row, having the index zero series as the dependente variable in relation with all the others.
-The \verb"'all-full'" option is operated by the function \verb"dmc_of_all_as_y".
+The \verb"'all-full'" option is operated by the function \verb"dmc_of_all_as_y()".
 
 There is a restriction to the \verb"dmcx2_of" array: the independent values must be in crescent order.
-There is a check in the \verb"dmcx2" that raises an error if the values are out of order, and suggests the use of a utility function named \verb"ordering_x_dmcx2_of".
-The code below shows an example of a manually defined \verb"dmcx2_of" array with 2 elements out of order in the second row, the \verb"AssertionError" output, the way of fixing it with the \verb"ordering_x_dmcx2_of" function and finally the outputted ordered array.
+There is a check in the \verb"dmcx2" that raises an error if the values are out of order, and suggests the use of a utility function named \verb"ordering_x_dmcx2_of()".
+The code below shows an example of a manually defined \verb"dmcx2_of" array with 2 elements out of order in the second row, the \verb"AssertionError" output, the way of fixing it with the \verb"ordering_x_dmcx2_of()" function and finally the outputted ordered array.
 

### 2025-01-30 paper_v02/article.tex
paper version 02

@@
 
+\subsection{Implementation of the Detrended Cross-correlation Coefficient function}
+
 It´s important to understand the calculation steps, the role of the \verb"DCCA_of" array and how it fits in the goals of the package implementation.
@@
 
-After all the pairs are calculated, the algorithm goes back to the
+After all the pairs are calculated, the algorithm goes back to the each box \verb"for" loop and head to the next box for the current time scale.
+when all the all the boxes are calculated, the final results for the $DFA$, $DCCA$ and $\rho_{DCCA}$ are evaluated and saved in the output vectors \verb"F_DFA_arr", \verb"DCCA_arr" and \verb"P_DCCA_arr" respectively.
 
@@
 
-The \verb"dmcx2" function runs the \verb"p_dcca()" with \verb"P_DCCA_output_matrix" set to \verb"True".
+\subsection{Implementation of the Detrended Multiple Cross-correlation Coefficient function}
+
+The \verb"dmcx2()" function runs the \verb"p_dcca()" with \verb"P_DCCA_output_matrix" set to \verb"True" in the background.
 There is no \verb"DCCA_of" input for the $DMC_x^2$ function, instead there is a \verb"dmcx2_of" parameter.
 This input receives a 2D matrix where each line represents the indexes of the series to be used in Equation \ref{eq:dmcx2}.
-The first value of each row is the index of the serie used as the dependent variable, the others, the independent ones.
+For each row, the first elements is the index of the serie used as the dependent variable, the others, the index of the independent ones.
 There are two literal strings that can be used, for convenience as inputs for this parameter: \verb"'all-full'", that generates a 2D array with every series as the dependent variable against all the others; and \verb"'first-full'", with only one row, having the index zero series as the dependente variable in relation with all the others.
-The \verb"'all-full'" option is operated by the function \verb"dmc_of_all_as_y()".
+The \verb"'all-full'" option is conducted by calling \verb"dmc_of_all_as_y()", also available as a utility function.
 
@@
 \end{CodeChunk}
+
+With a given \verb"dmcx2_of" an array with all the necessary pairs for the $DCCA$ and $\rho_{DCCA}$ is automatically generated and used for the background \verb"p_dcca()" function to calculate da matrix as in Eq.
+\ref{eq:p_dcca_matrix}.
+The \verb"dcca_of_from_dmcx2_of()", also can be used as an utility function, receives the  \verb"dmcx2_of" as input and returns the \verb"dcca_of" array.
+With the matrix $\rho_{DCCA}$ assembled, two internal functions, that can also be used as utility functions, calculates the $DMc_{x}^2$ for all the lines in the \verb"dmcx2_of" matrix, for all the time scales.
+The first function, \verb"dmcx2_from_p_dcca_matrix()", that receives the $\rho_{DCCA}$ and the \verb"dmcx2_of" array, is presented in the code below.
+
+\begin{Code}
+def dmcx2_from_p_dcca_matrix(P_DCCA_arr: NDArray[np.float64], 
+      dmcx2_of: NDArray[np.float64]) -> NDArray[np.float64]:   
+    # DMCx2 output matrix
+    DMCx2_arr = np.full(shape=(P_DCCA_arr.shape[2], dmcx2_of.shape[0]),
+    fill_value=np.nan, dtype=P_DCCA_arr.dtype)
+
+    for n_index in range(P_DCCA_arr.shape[2]):
+        P_DCCA_arr_2D = P_DCCA_arr[:, :, n_index]
+        for j,  dmcx2_of_1D in enumerate(dmcx2_of):
+            DMCx2_arr[n_index, j] = dmcx2_from_p_dcca_matrix_2d(P_DCCA_arr_2D,
+                                                                dmcx2_of_1D)
+    return DMCx2_arr
+\end{Code}
+
+The \verb"dmcx2_from_p_dcca_matrix()" function executes a for loop over the $\rho_{DCCA}$ that separates the this 3D matrix in to the 2D matrices for each time scale.
+Nested in this loop, another \verb"for" extracts each line of the \verb"dmcx2_of" and passes, the extracted matrix and the line vector to the \verb"dmcx2_from_p_dcca_matrix_2d()", displayed here.
+
+\begin{Code}
+def dmcx2_from_p_dcca_matrix_2d(P_DCCA_arr_2D: NDArray[np.float64], 
+          dmcx2_of_1D: NDArray[np.float64]) -> NDArray[np.float64]:   
+  y_index = dmcx2_of_1D[0:1]
+  x_indexes = dmcx2_of_1D[1:]
+
+  mat_x = P_DCCA_arr_2D[np.ix_(x_indexes, x_indexes)]
+  vec_y = P_DCCA_arr_2D[np.ix_(x_indexes, y_index)]
+
+  return vec_y.T @ np.linalg.inv(mat_x) @ vec_y
+\end{Code}
+
+The \verb"dmcx2_from_p_dcca_matrix_2d()" function uses \pkg{Numpy} methods to preparate the data to apply Eq.
+\ref{eq:dmcx2}.
+The list of indexes is divided in \verb"y_index", holding an one item array wiht the index of the dependent variabel time series, and \verb"x_indexes", containing the indexes of the independent ones.
+The \pkg{Numpy} function \verb"np.ix_()" constructs index arrays that will index the cross product from a serie of 1D arrays as inputs.
 

### 2025-01-31 paper_v02/article.tex
paper version 02

@@
 \ref{eq:dmcx2}.
-The list of indexes is divided in \verb"y_index", holding an one item array wiht the index of the dependent variabel time series, and \verb"x_indexes", containing the indexes of the independent ones.
-The \pkg{Numpy} function \verb"np.ix_()" constructs index arrays that will index the cross product from a serie of 1D arrays as inputs.
+The list of indexes is divided in \verb"y_index", holding an one item array whit the index of the dependent variabel time series, and \verb"x_indexes", containing the indexes of the independent ones.
+The \pkg{Numpy} \verb"np.ix_()", althow it's not a very known funcntion of the library constructs index arrays that will use the cross product from a serie of 1D arrays as inputs.
+It's a very convenient way to extract a submatrix and a vector from the $\rho_{DCCA}$ matrix.
+The matrix, as assembled by the \verb"p_dcca()" function, will always need to have extracions of a 2D matrix and a vector, as we can see in Eq.~\ref{eq:dmcx2}.
+The code below is an example of the extraction process.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+import numpy as np
+arr = np.array([[1,2,3,4],
+                [2,1,5,6],
+                [3,5,1,7],
+                [4,6,7,1]])
+print(arr)
+\end{CodeInput}
+
+\begin{CodeOutput}
+[[1 2 3 4]
+ [2 1 5 6]
+ [3 5 1 7]
+ [4 6 7 1]]
+\end{CodeOutput}
+An illustrative \verb"arr" matrix is defined as $4 \times 4$ with all ones in the main diagonal and symmetric integer values in the other cells.
+Since the $\rho_{DCCA}$ ranges from $-1$ to $1$, those values should be interpreted as place holders.
+\begin{CodeInput}
+sub_mat_index = np.ix_([1,2,3], [1,2,3])
+print("index combination:\n", sub_mat_index)
+print("extracted matrix:\n",arr[sub_mat_index])
+\end{CodeInput}
+
+\begin{CodeOutput}
+index combination:
+(array([[1],
+      [2],
+      [3]]), array([[1, 2, 3]]))
+extracted matrix:
+[[1 5 6]
+[5 1 7]
+[6 7 1]]
+\end{CodeOutput}
+
+Above, a submatrix, holding the positions $1$ to $3$ is extracted using \verb"np.ix_()" function, and below, the extraction of the vector, first as a line and then as a column vector.
+
+\begin{CodeInput}
+sub_mat_vec = np.ix_([0], [1,2,3])
+print("line vectror:\n", arr[sub_mat_vec])
+sub_mat_vec = np.ix_([1,2,3],[0])
+print("column vectror:\n", arr[sub_mat_vec])
+\end{CodeInput}
+
+\begin{CodeOutput}
+line vectror:
+[[2 3 4]]
+column vectror:
+[[2]
+[3]
+[4]]
+\end{CodeOutput}
+\end{CodeChunk}
+
+In the \pkg{Zebende} package, the vector is extracted as a column for better coherence with the $DMC_{x}^{2}$ theory.
+The method also works for dependent variable different of index $0$.
+The resulting matrix will preserve the diagonal as $1$, the symmetry regarding the main diagonal and the order of elements in the column vector respeted.
+the code below extract the index $1$ series as the dependent and the others as independent.
+\begin{CodeChunk}
+\begin{CodeInput}
+sub_mat_index = np.ix_([0,2,3], [0,2,3])
+print("index combination:\n", sub_mat_index)
+print("extracted matrix:\n",arr[sub_mat_index])
+sub_mat_vec = np.ix_([0,2,3],[1])
+print("column vectror:\n", arr[sub_mat_vec])
+\end{CodeInput}
+
+\begin{CodeOutput}
+index combination:
+(array([[0],
+      [2],
+      [3]]), array([[0, 2, 3]]))
+extracted matrix:
+[[1 3 4]
+[3 1 7]
+[4 7 1]]
+column vectror:
+[[2]
+[5]
+[6]]
+\end{CodeOutput}
+\end{CodeChunk}
 

### 2025-02-05 paper_v02/article.tex
paper version 02

@@
 Comparing the algorithms, the first three are basically identical, the only difference is that the $DCCA$ method apply those steps to two series.
-The step four of the $DFA$ can be considered an analogous of the variance, replacing the average subtraction (in the variance) for the values obtained by the polinomial fit(estimated seires); and the equivalent step of the $DCCA$ is, in the same terms, compared to the covariance between the two series.
+The step four of the $DFA$ can be considered an analogous of the variance, replacing the average subtraction (in the variance) for the values obtained by the polynomial fit (estimated series); and the equivalent step of the $DCCA$ is, in the same terms, compared to the covariance between the two series.
 The technique of fitting a curve, interpreted as a trend inside each box, and subtracting the value estimated by the trend from the actual value in the integrated series $(X_{k}-\widetilde{X}_{k})$ from now on will be called \textbf{detrended values}($DV$).
@@
 
-The implementation of the \pkg{Zebende} package follows some well defined golas:
+The implementation of the \pkg{Zebende} package follows some well defined goals:
 
@@
 The \proglang{Python} language was chosen because it´s one of the most used languages in the data analyses field and have a great support for statistical tools and machine learning algorithms.
-The are a pletora of tools to load and manipulate data (\pkg{Pandas}, \pkg{Polar}, \pkg{PySpark} ...), execute statistical analyzes (\pkg{Numpy}, \pkg{SciPy}, \pkg{StatsModels}~...), machine learning (\pkg{Pytorch}, \pkg{TensorFlow}, \pkg{Scykit Learn}~...) and data visualization(\pkg{Matplotlib}, \pkg{Seaborn}, \pkg{Plotly}~...) among other data related applications.
+The are a plethora of tools to load and manipulate data (\pkg{Pandas}, \pkg{Polar}, \pkg{PySpark} ...), execute statistical analyzes (\pkg{Numpy}, \pkg{SciPy}, \pkg{StatsModels}~...), machine learning (\pkg{Pytorch}, \pkg{TensorFlow}, \pkg{Scykit Learn}~...) and data visualization(\pkg{Matplotlib}, \pkg{Seaborn}, \pkg{Plotly}~...) among other data related applications.
 
@@
 For each row, the first elements is the index of the serie used as the dependent variable, the others, the index of the independent ones.
-There are two literal strings that can be used, for convenience as inputs for this parameter: \verb"'all-full'", that generates a 2D array with every series as the dependent variable against all the others; and \verb"'first-full'", with only one row, having the index zero series as the dependente variable in relation with all the others.
+There are two literal strings that can be used, for convenience as inputs for this parameter: \verb"'all-full'", that generates a 2D array with every series as the dependent variable against all the others; and \verb"'first-full'", with only one row, having the index zero series as the dependent variable in relation with all the others.
 The \verb"'all-full'" option is conducted by calling \verb"dmc_of_all_as_y()", also available as a utility function.
-
-There is a restriction to the \verb"dmcx2_of" array: the independent values must be in crescent order.
-There is a check in the \verb"dmcx2" that raises an error if the values are out of order, and suggests the use of a utility function named \verb"ordering_x_dmcx2_of()".
-The code below shows an example of a manually defined \verb"dmcx2_of" array with 2 elements out of order in the second row, the \verb"AssertionError" output, the way of fixing it with the \verb"ordering_x_dmcx2_of()" function and finally the outputted ordered array.
-
-\begin{CodeChunk}
-  \begin{CodeInput}
-dmcx2_of = np.array([ [0, 1, 2, 3],
-                      [1, 0, 3, 2],
-                      [2, 0, 1, 3],
-                      [3, 0, 1, 2]  ])
-
-dfa , dcca, pdcca, dmc = zb.dmcx2(int_data, tws, dmcx2_of=dmcx2_of)
-  \end{CodeInput}
-  \begin{CodeOutput}
-AssertionError: 
-dmcx2_of x values out of order: 
-use zebende.ordering_x_dmcx2_of(dmcx2_of)
- to fix it before passing the dmcx2_of values to zebende.dmcx2() function"
-  \end{CodeOutput}
-  \begin{CodeInput}
-dmcx2_of = ordering_x_dmcx2_of(dmcx2_of)
-print(dmcx2_of)
-  \end{CodeInput}
-  \begin{CodeOutput}
-[[0 1 2 3]
- [1 0 2 3]
- [2 0 1 3]
- [3 0 1 2]]
-  \end{CodeOutput}
-\end{CodeChunk}
 
@@
 \end{CodeOutput}
+\end{CodeChunk}
 An illustrative \verb"arr" matrix is defined as $4 \times 4$ with all ones in the main diagonal and symmetric integer values in the other cells.
 Since the $\rho_{DCCA}$ ranges from $-1$ to $1$, those values should be interpreted as place holders.
+\begin{CodeChunk}
 \begin{CodeInput}
@@
 \end{CodeOutput}
+\end{CodeChunk}
 
 Above, a submatrix, holding the positions $1$ to $3$ is extracted using \verb"np.ix_()" function, and below, the extraction of the vector, first as a line and then as a column vector.
+\begin{CodeChunk}
 
@@
 \end{CodeChunk}
+
+The idea of separating the calculations of the $\rho{DCCA}$ calculatons in three distinct functons aims to different workflows.
+The \verb"dmcx2()" function calculates and outputs all the prerequisites, as the $DFA$, $DCCA$, $\rho_{DCCA}$ along with the $DMC_{x}^{2}$.
+This is the most practical way of getting all this calculations conducted.
+But the task can also be diveded in two: first use the \verb"p_dcca()" function to generate $DFA$, $DCCA$, $\rho_{DCCA}$ outputs, analyze the outputs and then get the $DMC_{x}^{2}$ using \verb"dmcx2_from_p_dcca_matrix()".
+
+Function \verb"dmcx2_from_p_dcca_matrix_2d()" can be used to more customizable applications.
+Imagine a usecase where only the $\rho_{DCCA}$ anti-correlation, above a certain trashhold, in relation to the dependent variable, with the \verb"dmcx2_0f" set to \verb"all-full".
+From previous $\rho_{DCCA}$ studies is known that this coefficient can vary from posite to negative and vice  versa in different time scales.
+This implies that the $\rho_{DCCA}$ matrix should be analyzed in every time scale from every line of the \verb"dmcx2_of", also implies that the \verb"dmcx2_of" may not be a matrix in the sence that the rows may have different.
+The \pkg{Numpy} ND Array could not hold that.
+Many different workarounds could be proposed for that situation.
+In the implementation of this package, function \verb"dmcx2_from_p_dcca_matrix_2d()" allow the user to make a costum code that extracts the  $\rho_{DCCA}$ matrix for the current stime scale and extracts the elements that fit the rules from each \verb"dmcx2_of" row.
+
+\subsection{Zig implementation}
+
+The \proglang{Python} implementation successfully reflects the implementation goals presented in Sec.~\ref{sec:optimization} but the performance could be
+
+Zig language is one low-level language that gain popularity in recent years~\citep{10820804}.
+
+\subsection{Algorithm optimization}
 

### 2025-02-09 paper_v02/article.tex
paper version 02

@@
 
-The \proglang{Python} implementation successfully reflects the implementation goals presented in Sec.~\ref{sec:optimization} but the performance could be
+The \proglang{Python} implementation successfully reflects the implementation goals presented in Sec.~\ref{sec:optimization} but the performance could benefit with the integration of a low-level language.
+\proglang{Zig} was choosen to enhance the algorithms perpormance.
+It's a low-level language that gain popularity in recent years and provides performance similar to \proglang{C} and proglang{Fortran} in some scenarios~\citep{10820804}.
 
-Zig language is one low-level language that gain popularity in recent years~\citep{10820804}.
+The interest for using \proglang{ZIg} for this project also relies on the cross-compiling capabilities of the \proglang{Zig} compiler.
+For a small research group mantaining a package could be chalenging and the ability to to compile all releases for all platforms in a single machine is a great advantage.
+The \proglang{Zig} compiler can generate binaries for Windows, Linux and MacOS from a single machine.
+The \proglang{Zig} compiler is also very fast, and the language is very easy to learn, with a syntax that is very similar to \proglang{C}.
+
+The implementaton focus on writing technics $\rho_{DCCA}$ function in \proglang{Zig} exposed as a \proglang{C} Application Binary Interface (ABI), called by \proglang{Python} using the \pkg{ctypes} package.
+The $\rho_{DCCA}$ function is the most computational expensive part of the package, and the performance gain was expected to be significant.
+
+The output arrays are allocated in the \proglang{Python} side and passed as pointers to the \proglang{Zig} function thogheter with the series matrix, \verb"tws" and \verb"DCCA_of" arrays.
+The \proglang{Zig} function receives the pointers and the size of the arrays as inputs, and the results are stored in the same memory space.
+Before passing to \proglang{Zig} must be assured that the arrays are contiguous, using the \verb"np.ascontiguousarray()" function.
+The boolean parameter \verb"P_dcca_matrix_output" is also passed to the \proglang{Zig} function, to determine if the output of the $\rho_{DCCA}$ should be a table or a matrix.
+
+The \proglang{Zig} implementation folows the same steps as the \proglang{Python} one, but repspecting languages differences.
+In \proglang{Python}, using \pkg{Numpy}, all the time series calculations occour in the same line of code, using the \pkg{Numpy} broadcasting capabilities.
+In \proglang{Zig}, the calculations are made in a nested \verb"for" loop.
 
 \subsection{Algorithm optimization}
+
+Althow the \proglang{Zig} implementation presents a significant performance gain,the algorithm can be optimized.
+The strategie is to focus on repeated calculatins in the process.
+The code below shows the calculations of the polynomial fit before the optimizarion.
+
+\begin{Code}
+  pub fn lin_ls_fit(win: []f64, time: []f64) [2]f64 {
+    var x_sum: f64 = 0;
+    var y_sum: f64 = 0;
+    var xy_sum: f64 = 0;
+    var x2_sum: f64 = 0;
+    for (win, time) |w, t| {
+        x_sum += t;
+        y_sum += w;
+        xy_sum += t * w;
+        x2_sum += pow(f64, t, 2);
+    }
+    const n: f64 = @as(f64, @floatFromInt(time.len));
+    //slope
+    const slope: f64 = (((n * xy_sum) - (x_sum * y_sum)) /
+        ((n * x2_sum) - (pow(f64, x_sum, 2))));
+    //inter
+    const inter: f64 = ((y_sum - (slope * x_sum)) /
+        (n));
+    //result
+    return [_]f64{ slope, inter };
+}
+\end{Code}
+
+The code above calculates the slope and the intercept of a linear least squares fit.
+The function runs on every box for every time scale.
+For consecutive boxes, the value of the sums from the previous box is used to calculate the next one widhout a loop by implementing a simple technic: when a box is calculated, the first value is saved in a temporary variable, to calculate the next sum, the previous sum is suntracted from the temporary variable and the new value is added.
+Also, to optimize the calculations from one time scale to the next, the first sums for the first box are saved in a temporary variable.
+Considering the current time scale as $tws_prev$ and the consecutive as $tws_current$.
+To calculate the current value, the algorithm takes the sums from the $tws_prev$ and add the values with indexes that exceed the $tws_prev$ size.
+The code was also rewritten with structs, to hold the temporary values and the functons for better readability.
+The code below shows the optimized version of polynomial fitting calculations.
+
+\begin{Code}
+
+fn shiftWindow( self: *MainOperator,
+                n: usize, win_start: usize,
+                F_DFA_ptr: *allowzero [*c]f64) void {
+    self.time_window = self.time[win_start ..][ .. (n + 1)];
+    // print("win_start {}\n", .{win_start});
+    if (win_start != 0) {
+        // updating sum_x
+        self.current.sum_x = self.current.sum_x - self.left_x + self.time_window[n];
+        // updating sum x^2
+        self.current.sum_x2 = self.current.sum_x2 - 
+        pow(f64, self.left_x, 2) + pow(f64, self.time_window[n], 2);
+        // updating y and y*x for every serie
+        for (self.series, 0..) |serie, sr_index| {
+            serie.current.sum_y = serie.current.sum_y - 
+            serie.left_y + serie.serie[win_start + n];
+
+            serie.current.sum_xy = serie.current.sum_xy - 
+            (self.left_x * serie.left_y) + 
+            (self.time_window[n] * serie.serie[win_start + n]);
+            serie.left_y = serie.serie[win_start];
+            self.detrended(serie, win_start, &F_DFA_ptr.*[sr_index]);
+        }
+    } else { // win_start == 0
+        self.current.window_len = n + 1;
+
+        for (self.previous.window_len..self.current.window_len) |i| {
+            self.previous.sum_x += self.time_window[i];
+            self.previous.sum_x2 += pow(f64, self.time_window[i], 2);
+
+            for (self.series) |serie| {
+                serie.previous.sum_y += serie.serie[i];
+                serie.previous.sum_xy += self.time_window[i] * serie.serie[i];
+            }
+        }
+
+        // updating current sum values
+        self.current.sum_x = self.previous.sum_x;
+        self.current.sum_x2 = self.previous.sum_x2;
+
+        for (self.series, 0..) |serie, sr_index| {
+            serie.current.sum_xy = serie.previous.sum_xy;
+            serie.current.sum_y = serie.previous.sum_y;
+
+            serie.left_y = serie.serie[win_start];
+            self.detrended(serie, win_start, &F_DFA_ptr.*[sr_index]);
+        }
+    }
+    self.left_x = self.time_window[0];
+}
+\end{Code}
+In the code above, the \verb"MainOperator" is a structure that holds the time series(y), the time window, the sums and the temporary values.
+The \verb"shiftWindow()" function is a method of the \verb"MainOperator" structure.
+The function receives the size of the box, the index of the first element of the box, and a pointer to the $F_{DFA}^{2}$ array.
+The function is called for every box in every time scale.
+The function is responsible for updating the sums of the time series, the sums of the $DV$ and the $DV$ matrix.
+
+The function also calls the \verb"detrended()" function, that calculates the $DV$ for each time series.
+The function is called for every time series in the \verb"MainOperator" structure.
+The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
+The function is called for every time series in the \verb"MainOperator" structure.
+The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
+The function is called for every time series in the \verb"MainOperator" structure.
+The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
+The function is called for every time series in the \verb"MainOperator" structure.
+The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
+The function is called for every time series in the \verb"MainOperator" structure.
+The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
+The function is called for every time series in the \verb"MainOperator" structure.
+The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
+The function is called for every time series in the \verb"MainOperator" structure.
+The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series
+
+The three implementations, pure \proglang{Python}, \proglang{ZIg} and optimized \proglang{Zig}, where tested and the results are presented in the next section.
 

### 2025-02-14 paper_v02/article.tex
paper version 02

@@
   \item for all the boxes of a time scale, the $DCCA$ is calculated as:\\[10pt]
-        $F_{DCCA}^{2}(n) =\frac{1}{N - n} \sum_{i=1}^{N-n} f_{DCCA}^{2}(n, i)$;
-  \item for a number of different timescales ($n$), with possible values
-        $4 \le n \le \frac{N}{4}$ the $F_{DCCA}^{2}$ function is calculated to find a relation among
-        $F_{DCCA}^{2} \times n$
 \end{enumerate}
@@
 
-The function also calls the \verb"detrended()" function, that calculates the $DV$ for each time series.
-The function is called for every time series in the \verb"MainOperator" structure.
-The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
-The function is called for every time series in the \verb"MainOperator" structure.
-The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
-The function is called for every time series in the \verb"MainOperator" structure.
-The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
-The function is called for every time series in the \verb"MainOperator" structure.
-The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
-The function is called for every time series in the \verb"MainOperator" structure.
-The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
-The function is called for every time series in the \verb"MainOperator" structure.
-The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series.
-The function is called for every time series in the \verb"MainOperator" structure.
-The \verb"detrended()" function is not shown in the code above, but it is a simple function that calculates the $DV$ for a time series
+\begin{equation}
+\sum_{k=i}^{i + n}S_k = \sum_{j=i-1}^{n}S_j - S_{i-1} + S_{n+1}
+\end{equation}
 

### 2025-02-20 paper_v02/article.tex
paper version 02

@@
 The Detrended Cross-correlation Coefficient ($\rho_{DCCA}$)~\citep{Zebende2011} is a widely used coefficient that measures the cross-correlation between two non-stationary time series.
-It´s an extension of the Detrended Fluctuation Analysis ($DFA$)~\citep{Peng_1994} and the Detrended Cross-correlation Analysis ($DCCA$)~\citep{Podobnik2008}: while the $DFA$ calculates the self-affinity and long-memory properties of a time series data, and the $DCCA$ analyses power-law cross correlations between two different non-stationarity time series, the $\rho_{DCCA}$ coefficient quantifies this cross-correlation in simple values ranging from $-1$ to $1$, where $-1$ indicates a perfect anti-correlation between the series, $1$ a perfect correlation and zero ($0$) no correlation at all.
+It´s an extension of the Detrended Fluctuation Analysis ($DFA$)~\citep{Peng_1994} and the Detrended Cross-correlation Analysis ($DCCA$)~\citep{Podobnik2008}: while the $DFA$ calculates the self-affinity and long-memory properties of a time series data, and the $DCCA$ analyses power-law cross correlations between two different non-stationary time series, the $\rho_{DCCA}$ coefficient quantifies this cross-correlation in simple values ranging from $-1$ to $1$, where $-1$ indicates a perfect anti-correlation between the series, $1$ a perfect correlation and zero ($0$) no correlation at all.
 
@@
 The \verb"detrended_mat" array has a number of lines equal to the number of points in a time scale box ($n +1$, resized for each $n$) and column count also equal to the number of time series (same size in each $n$).
-The first two tempoaray arrays will store data for all the boxes in the time scale, the last one will be used in each box and will have the values replaced in the next one, until the last box of the time scale.
+The first two temporary arrays will store data for all the boxes in the time scale, the last one will be used in each box and will have the values replaced in the next one, until the last box of the time scale.
 Than all the arrays will be cleared and recreated with shapes calculated with the new value of $n$.
@@
 After the allocation of the temporary arrays, the second \verb"for" loop operates in every box for a certain time scale $n$.
-In each box, the \verb"detrended_series()" function execute a one degree polynomial fit, subtract the values of the serie with the value given by the interpolated curve($DV$) and stores this values in the \verb"detrended_mat".
+In each box, the \verb"detrended_series()" function execute a one degree polynomial fit, subtract the values of the series with the value given by the interpolated curve($DV$) and stores this values in the \verb"detrended_mat".
 After that, the \textbf{mean of the square of each $DV$ in the box} ($f_{DFA}^{2}(n, i)$) is calculated in the current box for every time series and stored in the \verb"f2dfa_n" array in the line associated with the current box, and the column related to each time series.
@@
 This input receives a 2D matrix where each line represents the indexes of the series to be used in Equation \ref{eq:dmcx2}.
-For each row, the first elements is the index of the serie used as the dependent variable, the others, the index of the independent ones.
+For each row, the first elements is the index of the series used as the dependent variable, the others, the index of the independent ones.
 There are two literal strings that can be used, for convenience as inputs for this parameter: \verb"'all-full'", that generates a 2D array with every series as the dependent variable against all the others; and \verb"'first-full'", with only one row, having the index zero series as the dependent variable in relation with all the others.
@@
 
-The \verb"dmcx2_from_p_dcca_matrix_2d()" function uses \pkg{Numpy} methods to preparate the data to apply Eq.
+The \verb"dmcx2_from_p_dcca_matrix_2d()" function uses \pkg{Numpy} methods to prepare the data to apply Eq.
 \ref{eq:dmcx2}.
-The list of indexes is divided in \verb"y_index", holding an one item array whit the index of the dependent variabel time series, and \verb"x_indexes", containing the indexes of the independent ones.
-The \pkg{Numpy} \verb"np.ix_()", althow it's not a very known funcntion of the library constructs index arrays that will use the cross product from a serie of 1D arrays as inputs.
-It's a very convenient way to extract a submatrix and a vector from the $\rho_{DCCA}$ matrix.
-The matrix, as assembled by the \verb"p_dcca()" function, will always need to have extracions of a 2D matrix and a vector, as we can see in Eq.~\ref{eq:dmcx2}.
+The list of indexes is divided in \verb"y_index", holding an one item array whit the index of the dependent variable time series, and \verb"x_indexes", containing the indexes of the independent ones.
+The \pkg{Numpy} \verb"np.ix_()", although  it's not a very known function of the library constructs index arrays that will use the cross product from a series of 1D arrays as inputs.
+It's a very convenient way to extract a sub matrix and a vector from the $\rho_{DCCA}$ matrix.
+The matrix, as assembled by the \verb"p_dcca()" function, will always need to have extractions of a 2D matrix and a vector, as we can see in Eq.~\ref{eq:dmcx2}.
 The code below is an example of the extraction process.
@@
 
-Above, a submatrix, holding the positions $1$ to $3$ is extracted using \verb"np.ix_()" function, and below, the extraction of the vector, first as a line and then as a column vector.
+Above, a sub matrix, holding the positions $1$ to $3$ is extracted using \verb"np.ix_()" function, and below, the extraction of the vector, first as a line and then as a column vector.
 \begin{CodeChunk}
@@
 sub_mat_vec = np.ix_([0], [1,2,3])
-print("line vectror:\n", arr[sub_mat_vec])
+print("line vector:\n", arr[sub_mat_vec])
 sub_mat_vec = np.ix_([1,2,3],[0])
-print("column vectror:\n", arr[sub_mat_vec])
+print("column vector:\n", arr[sub_mat_vec])
 \end{CodeInput}
@@
 \begin{CodeOutput}
-line vectror:
+line vector:
 [[2 3 4]]
-column vectror:
+column vector:
 [[2]
@@
 The method also works for dependent variable different of index $0$.
-The resulting matrix will preserve the diagonal as $1$, the symmetry regarding the main diagonal and the order of elements in the column vector respeted.
+The resulting matrix will preserve the diagonal as $1$, the symmetry regarding the main diagonal and the order of elements in the column vector respected.
 the code below extract the index $1$ series as the dependent and the others as independent.
@@
 sub_mat_vec = np.ix_([0,2,3],[1])
-print("column vectror:\n", arr[sub_mat_vec])
+print("column vector:\n", arr[sub_mat_vec])
 \end{CodeInput}
@@
 [4 7 1]]
-column vectror:
+column vector:
 [[2]
@@
 
-The idea of separating the calculations of the $\rho{DCCA}$ calculatons in three distinct functons aims to different workflows.
+The idea of separating the calculations of the $\rho{DCCA}$ calculations in three distinct functions aims to different workflows.
 The \verb"dmcx2()" function calculates and outputs all the prerequisites, as the $DFA$, $DCCA$, $\rho_{DCCA}$ along with the $DMC_{x}^{2}$.
 This is the most practical way of getting all this calculations conducted.
-But the task can also be diveded in two: first use the \verb"p_dcca()" function to generate $DFA$, $DCCA$, $\rho_{DCCA}$ outputs, analyze the outputs and then get the $DMC_{x}^{2}$ using \verb"dmcx2_from_p_dcca_matrix()".
+But the task can also be divided in two: first use the \verb"p_dcca()" function to generate $DFA$, $DCCA$, $\rho_{DCCA}$ outputs, analyze the outputs and then get the $DMC_{x}^{2}$ using \verb"dmcx2_from_p_dcca_matrix()".
 
 Function \verb"dmcx2_from_p_dcca_matrix_2d()" can be used to more customizable applications.
-Imagine a usecase where only the $\rho_{DCCA}$ anti-correlation, above a certain trashhold, in relation to the dependent variable, with the \verb"dmcx2_0f" set to \verb"all-full".
-From previous $\rho_{DCCA}$ studies is known that this coefficient can vary from posite to negative and vice  versa in different time scales.
-This implies that the $\rho_{DCCA}$ matrix should be analyzed in every time scale from every line of the \verb"dmcx2_of", also implies that the \verb"dmcx2_of" may not be a matrix in the sence that the rows may have different.
+Imagine a use case where only the $\rho_{DCCA}$ anti-correlation, inside a certain range, in relation to the dependent variable, with the \verb"dmcx2_0f" set to \verb"all-full".
+From previous $\rho_{DCCA}$ studies is known that this coefficient can vary from positive to negative and vice  versa in different time scales.
+This implies that the $\rho_{DCCA}$ matrix should be analyzed in every time scale from every line of the \verb"dmcx2_of", also implies that the \verb"dmcx2_of" may not be a matrix in the since that the rows may have different.
 The \pkg{Numpy} ND Array could not hold that.
 Many different workarounds could be proposed for that situation.
-In the implementation of this package, function \verb"dmcx2_from_p_dcca_matrix_2d()" allow the user to make a costum code that extracts the  $\rho_{DCCA}$ matrix for the current stime scale and extracts the elements that fit the rules from each \verb"dmcx2_of" row.
+In the implementation of this package, function \verb"dmcx2_from_p_dcca_matrix_2d()" allow the user to make a custom code that extracts the  $\rho_{DCCA}$ matrix for the current time scale and extracts the elements that fit the rules from each \verb"dmcx2_of" row.
 
@@
 The \proglang{Python} implementation successfully reflects the implementation goals presented in Sec.~\ref{sec:optimization} but the performance could benefit with the integration of a low-level language.
-\proglang{Zig} was choosen to enhance the algorithms perpormance.
-It's a low-level language that gain popularity in recent years and provides performance similar to \proglang{C} and proglang{Fortran} in some scenarios~\citep{10820804}.
+\proglang{Zig} was chosen to enhance the algorithms performance.
+It's a low-level language that gain popularity in recent years and provides performance similar to \proglang{C} and \proglang{Fortran} in some scenarios~\citep{10820804}.
 
 The interest for using \proglang{ZIg} for this project also relies on the cross-compiling capabilities of the \proglang{Zig} compiler.
-For a small research group mantaining a package could be chalenging and the ability to to compile all releases for all platforms in a single machine is a great advantage.
+For a small research group maintaining a package could be challenging and the ability to to compile all releases for all platforms in a single machine is a great advantage.
 The \proglang{Zig} compiler can generate binaries for Windows, Linux and MacOS from a single machine.
@@
 
-The implementaton focus on writing technics $\rho_{DCCA}$ function in \proglang{Zig} exposed as a \proglang{C} Application Binary Interface (ABI), called by \proglang{Python} using the \pkg{ctypes} package.
+The implementation focus on writing technics $\rho_{DCCA}$ function in \proglang{Zig} exposed as a \proglang{C} Application Binary Interface (ABI), called by \proglang{Python} using the \pkg{ctypes} package.
 The $\rho_{DCCA}$ function is the most computational expensive part of the package, and the performance gain was expected to be significant.
 
-The output arrays are allocated in the \proglang{Python} side and passed as pointers to the \proglang{Zig} function thogheter with the series matrix, \verb"tws" and \verb"DCCA_of" arrays.
+The output arrays are allocated in the \proglang{Python} side and passed as pointers to the \proglang{Zig} function together with the series matrix, \verb"tws" and \verb"DCCA_of" arrays.
 The \proglang{Zig} function receives the pointers and the size of the arrays as inputs, and the results are stored in the same memory space.
@@
 
-The \proglang{Zig} implementation folows the same steps as the \proglang{Python} one, but repspecting languages differences.
-In \proglang{Python}, using \pkg{Numpy}, all the time series calculations occour in the same line of code, using the \pkg{Numpy} broadcasting capabilities.
+The \proglang{Zig} implementation follows the same steps as the \proglang{Python} one, but repspecting languages differences.
+In \proglang{Python}, using \pkg{Numpy}, all the time series calculations occur in the same line of code, using the \pkg{Numpy} broadcasting capabilities.
 In \proglang{Zig}, the calculations are made in a nested \verb"for" loop.
@@
 
-Althow the \proglang{Zig} implementation presents a significant performance gain,the algorithm can be optimized.
-The strategie is to focus on repeated calculatins in the process.
-The code below shows the calculations of the polynomial fit before the optimizarion.
+Although the \proglang{Zig} implementation presents a significant performance gain, the algorithm still can be optimized.
+The strategy is to focus on avoiding repeated calculations in the process.
+The code below shows the calculations of the polynomial fit before the optimization.
 
@@
 The code above calculates the slope and the intercept of a linear least squares fit.
-The function runs on every box for every time scale.
-For consecutive boxes, the value of the sums from the previous box is used to calculate the next one widhout a loop by implementing a simple technic: when a box is calculated, the first value is saved in a temporary variable, to calculate the next sum, the previous sum is suntracted from the temporary variable and the new value is added.
+The function runs on every box for every time scale.In the optimized version, for consecutive boxes, the value of the sums from the previous box is used to calculate the next one without a loop for every item in the box.
+
+\begin{equation}
+  \label{eq:sum_opt}
+  \forall~1<i\leq(N-n),~\sum_{k=i}^{i + n}S_k = \left(\sum_{j=i-1}^{(i+n)-1}S_j\right) - S_{i-1} + S_{n+1}
+  \end{equation}
+
+The expression in Eq.~\ref{eq:sum_opt} stands that the sum of the values in a box is equal to the sum of the previous box minus the first value of the previous box plus the last value of the current box.
+According to that, the
+when a box is calculated, the sum and the first value is saved in temporary variables, to calculate the next sum, the previous sum is subtracted from the temporary variable and the new value is added.
+The bigger the box, the bigger the gain in performance.
+
 Also, to optimize the calculations from one time scale to the next, the first sums for the first box are saved in a temporary variable.
@@
 To calculate the current value, the algorithm takes the sums from the $tws_prev$ and add the values with indexes that exceed the $tws_prev$ size.
-The code was also rewritten with structs, to hold the temporary values and the functons for better readability.
+The code was also rewritten with structs, to hold the temporary values and the functions for better readability.
 The code below shows the optimized version of polynomial fitting calculations.
@@
 The function is responsible for updating the sums of the time series, the sums of the $DV$ and the $DV$ matrix.
-
-\begin{equation}
-\sum_{k=i}^{i + n}S_k = \sum_{j=i-1}^{n}S_j - S_{i-1} + S_{n+1}
-\end{equation}
 

### 2025-02-25 paper_v02/article.tex
Expressions

@@
 \begin{equation}
-  \label{eq:sum_opt}
-  \forall~1<i\leq(N-n),~\sum_{k=i}^{i + n}S_k = \left(\sum_{j=i-1}^{(i+n)-1}S_j\right) - S_{i-1} + S_{n+1}
+  \label{eq:sum_opt_x}
+  \forall~1<i\leq(N-n),~\sum_{k=i}^{i + n}T_k = \left(\sum_{j=i-1}^{(i+n)-1}T_j\right)~-~T_{i-1}~+~T_{i + n}
 \end{equation}
 
-The expression in Eq.~\ref{eq:sum_opt} stands that the sum of the values in a box is equal to the sum of the previous box minus the first value of the previous box plus the last value of the current box.
+\begin{equation}
+  \label{eq:sum_opt_x2}
+  \forall~1<i\leq(N-n),~\sum_{k=i}^{i + n}T_k^2 = \left(\sum_{j=i-1}^{(i+n)-1}T_j^2\right)~-~T_{i-1}^2~+~T_{i + n}^2
+\end{equation}
+
+\begin{equation}
+  \label{eq:sum_opt_y}
+  \forall~1<i\leq(N-n),~\sum_{k=i}^{i + n}S_k = \left(\sum_{j=i-1}^{(i+n)-1}S_j\right)~-~S_{i-1}~+~S_{i + n}
+\end{equation}
+
+\begin{equation}
+  \label{eq:sum_opt_xy}
+  \forall~1<i\leq(N-n),~\sum_{k=i}^{i + n} (S_k\times T_k) = \left(\sum_{j=i-1}^{(i+n)-1}(S_j \times T_j)\right)-(S_{i-1} \times T_{i-1})+(S_{i + n} \times T_{i + n})
+\end{equation}
+
+The expression in Eq.~\ref{eq:sum_opt_y} stands that the sum of the values in a box is equal to the sum of the previous box minus the first value of the previous box plus the last value of the current box.
 According to that, the

### 2025-02-27 paper_v02/article.tex
paper version 02

@@
 The code above calculates the slope and the intercept of a linear least squares fit.
-The function runs on every box for every time scale.In the optimized version, for consecutive boxes, the value of the sums from the previous box is used to calculate the next one without a loop for every item in the box.
+The function runs on every box for every time scale.
+In the optimized version, for consecutive boxes, the value of the sums from the previous box is used to calculate the next one without a loop for every item in the box.
+This technic proved to be very efficient in increasing performance.
+the expressions below show the effectiveness of the optimized algorithm.
 
@@
 
-The expression in Eq.~\ref{eq:sum_opt_y} stands that the sum of the values in a box is equal to the sum of the previous box minus the first value of the previous box plus the last value of the current box.
-According to that, the
-when a box is calculated, the sum and the first value is saved in temporary variables, to calculate the next sum, the previous sum is subtracted from the temporary variable and the new value is added.
-The bigger the box, the bigger the gain in performance.
+The expression in Eq.~\ref{eq:sum_opt_y} states that the sum of the values in a box is equal to the sum of the previous box minus the first value of the previous box plus the last value of the current box.
+According to that, when a box is calculated, the sum and the first value are saved in temporary variables.
+To calculate the next sum, the previous sum is subtracted from the temporary variable and the new value is added.
+The larger the box, the greater the performance gain.
 

### 2025-03-02 paper_v02/article.tex
paper version 02

@@
 
-The expression in Eq.~\ref{eq:sum_opt_y} states that the sum of the values in a box is equal to the sum of the previous box minus the first value of the previous box plus the last value of the current box.
-According to that, when a box is calculated, the sum and the first value are saved in temporary variables.
-To calculate the next sum, the previous sum is subtracted from the temporary variable and the new value is added.
-The larger the box, the greater the performance gain.
+The expressions (Eq.~\ref{eq:sum_opt_x},Eq.~\ref{eq:sum_opt_x2}, Eq.~\ref{eq:sum_opt_y} and Eq.\ref{eq:sum_opt_xy}) showcases how the optimizations can be implemented in the code.
+Each expression display that the sum of the values in a box can be calculated using the sum of the previous box and the values of the first and last elements of the current box.
+Eq.\ref{eq:sum_opt_x} presents the sum of the time stamps values and Eq.\ref{eq:sum_opt_x2} the sum of the squares of the time stamps values.
+This calculations occur once for every box in every time scale and is stored in variables.
+Eq.\ref{eq:sum_opt_y} and Eq.\ref{eq:sum_opt_xy} presents the same calculations for the series values.
+Storing values for each time series.
 
 Also, to optimize the calculations from one time scale to the next, the first sums for the first box are saved in a temporary variable.
-Considering the current time scale as $tws_prev$ and the consecutive as $tws_current$.
-To calculate the current value, the algorithm takes the sums from the $tws_prev$ and add the values with indexes that exceed the $tws_prev$ size.
+Considering the current time scale as \verb"tws_prev" and the consecutive as \verb"tws_current".
+To calculate the current value, the algorithm takes the sums from the \verb"tws_prev" and add the values with indexes that exceed the \verb"tws_prev" size.
 The code was also rewritten with structs, to hold the temporary values and the functions for better readability.
@@
 \end{Code}
-In the code above, the \verb"MainOperator" is a structure that holds the time series(y), the time window, the sums and the temporary values.
+In the code above, the \verb"MainOperator" is a structure that holds the time series($Y_n$), the time stamps(x), the sums for the time window and for each series and the temporary variables that holds the first values of the previous box.
 The \verb"shiftWindow()" function is a method of the \verb"MainOperator" structure.
@@
 The function is responsible for updating the sums of the time series, the sums of the $DV$ and the $DV$ matrix.
+
+The if statement in the code above is responsible for updating the sums for the current box. The first \verb"if" block is executed for
+every box except the first one in each time scale, where the \verb"else" block is executed instead.
+
+The \verb"if" block first update the values for the sum of the time stamps (\verb"sum_x") and the sum of the squares of the time stamps (\verb"sum_x2").
+Then, for every series, the sum of the series values (\verb"sum_y") and the sum of the product of the series and the time stamps are updated (\verb"sum_xy").
+Than the values are passed to the \verb"detrended" function that calculates The $DV$ and the $DFA$ for the current box.
+
+The \verb"else" block is executed for the first box in each time scale.
+For consecutive time scales, the values of the sums are updated using the values of the previous box, looping and adding only the values in the current box that don't overlap with the previous one.
+The \verb"detrended" function is also called for every series in the time scale.
 

### 2025-03-03 tests_r/test r_DCCA_PKG.R
r tests

@@ -1,0 +1,23 @@
+require("DCCA")
+
+tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
+         362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
+
+data <- read.csv("https://tinyurl.com/zbtestdata01")
+
+f3 <- data[, c("F3")]
+f6 <- data[, c("F6")]
+
+
+startTime <- Sys.time() 
+
+result =  rhodcca(f3, f6, m = tws, nu = 0, overlap = TRUE)
+
+endTime <- Sys.time()
+
+# prints recorded time 
+print(endTime - startTime)
+
+
+
+

### 2025-03-03 tests_r/test r_DFA_PKG.R

@@ -1,0 +1,19 @@
+require("DFA")
+
+tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
+        362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
+
+data <- read.csv("https://tinyurl.com/zbtestdata01")
+
+f3 <- data[, c("F3")]
+f6 <- data[, c("F6")]
+
+
+startTime <- Sys.time() 
+
+result =  rhoDCCA(f3, f6, scale ="F", box_size= tws)
+
+endTime <- Sys.time()
+
+# prints recorded time 
+print(endTime - startTime)

### 2025-03-03 tests_r/test r_DCCA_PKG.R
R tests and comparisson

@@ -4,11 +4,11 @@ tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104
          362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
 
 data <- read.csv("https://tinyurl.com/zbtestdata01")
-
+print(length(tws))
 f3 <- data[, c("F3")]
 f6 <- data[, c("F6")]
 
-
+print("Start")
 startTime <- Sys.time() 
 
 result =  rhodcca(f3, f6, m = tws, nu = 0, overlap = TRUE)

### 2025-03-03 tests_r/test r_DFA_PKG.R

@@ -1,17 +1,22 @@
 require("DFA")
 
+# tws scale
 tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
         362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
 
+
+print(length(tws))
+
+# read data
 data <- read.csv("https://tinyurl.com/zbtestdata01")
 
 f3 <- data[, c("F3")]
 f6 <- data[, c("F6")]
 
-
+print("Start")
 startTime <- Sys.time() 
 
-result =  rhoDCCA(f3, f6, scale ="F", box_size= tws)
+result_rhoDCCA =  rhoDCCA(f3, f6, scale ="F", box_size= tws)
 
 endTime <- Sys.time()
 

### 2025-03-03 tests_r/test r_DCCA_PKG.R
R tests and comparisson

@@ -1,21 +1,26 @@
 require("DCCA")
 
+# define time windows scales
 tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
          362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
 
+print("tws length",length(tws))
+
+# load data
 data <- read.csv("https://tinyurl.com/zbtestdata01")
-print(length(tws))
+
+# extract series
 f3 <- data[, c("F3")]
 f6 <- data[, c("F6")]
 
 print("Start")
 startTime <- Sys.time() 
 
-result =  rhodcca(f3, f6, m = tws, nu = 0, overlap = TRUE)
+result_DCCA_PKG =  rhodcca(f3, f6, m = tws, nu = 0, overlap = TRUE)
 
 endTime <- Sys.time()
 
-# prints recorded time 
+# print recorded time 
 print(endTime - startTime)
 
 

### 2025-03-03 tests_r/test r_DFA_PKG.R

@@ -1,24 +1,24 @@
 require("DFA")
 
-# tws scale
+# define time windows scales
 tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
         362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
 
-
-print(length(tws))
+print("tws length",length(tws))
 
 # read data
 data <- read.csv("https://tinyurl.com/zbtestdata01")
 
+# extract series
 f3 <- data[, c("F3")]
 f6 <- data[, c("F6")]
 
 print("Start")
 startTime <- Sys.time() 
 
-result_rhoDCCA =  rhoDCCA(f3, f6, scale ="F", box_size= tws)
+result_DFA_PKG =  rhoDCCA(f3, f6, scale ="F", box_size= tws)
 
 endTime <- Sys.time()
 
-# prints recorded time 
+# print recorded time 
 print(endTime - startTime)

### 2025-03-05 paper_v02/article.tex
tests

@@
 
+The algorithm was tested with data and results from the \citep{RIBEIRO2025107419}, the original data came from the \textit{Physionet} open-access website, available at this link:
+
+\begin{center}
+	\url{https://physionet.org/content/eegmmidb/1.0.0/}
+\end{center} 
+This comprehensive dataset, with 108 subjects processes for all the 12 experiments, with the $\rho_{DCCA}$ and $DMC_{x}^{2}$ calculated for all the subjects and experiments, where used to compare result and validate the implementations.
+
+Regarding performance, the pure \proglang{Python} implementation, the \proglang{Zig} implementation and the optimized \proglang{Zig} where compared with 3 oder packages that calculates the $\rho_{DCCA}$.
+Researching the \proglang{R} \textbf{CRAN} package repository, the \pkg{DCCA} \pkg{DFA} packages where found, and, from the \proglang{Python} \textbf{pypi}, the \pkg{DCCA} package was used.
+
+The \pkg{DFA} is a pure \proglang{R} implementation.
+The \pkg{DCCA} is a \proglang{R} package with a \proglang{Java} implementation.
+The \pkg{DCCA} is a \proglang{Python} implementation with massive use of \pkg{Numpy} broadcasting capabilities and memory access with strides to avoid loops.
+
+Regarding performance
+
 %% -- Summary/conclusions/discussion -------------------------------------------

### 2025-03-05 tests_r/test r_DCCA_PKG.R

@@ -2,17 +2,17 @@ require("DCCA")
 
 # define time windows scales
 tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
-         362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
+         362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570);
 
-print("tws length",length(tws))
+print(length(tws));
 
 # load data
 data <- read.csv("https://tinyurl.com/zbtestdata01")
 
 # extract series
-f3 <- data[, c("F3")]
-f6 <- data[, c("F6")]
-
+f3 <- data[, 1];
+f6 <- data[, c("F6")];
+print(f3)
 print("Start")
 startTime <- Sys.time() 
 
@@ -22,6 +22,8 @@ endTime <- Sys.time()
 
 # print recorded time 
 print(endTime - startTime)
+
+result_DCCA_PKG_pdcca = result_DCCA_PKG[["rhodcca"]]
 
 
 

### 2025-03-05 tests_r/test r_DFA_PKG.R

@@ -4,7 +4,7 @@ require("DFA")
 tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
         362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
 
-print("tws length",length(tws))
+print(length(tws))
 
 # read data
 data <- read.csv("https://tinyurl.com/zbtestdata01")
@@ -20,5 +20,6 @@ result_DFA_PKG =  rhoDCCA(f3, f6, scale ="F", box_size= tws)
 
 endTime <- Sys.time()
 
-# print recorded time 
+# print recorded time
+print("DFA_PKG result: ")
 print(endTime - startTime)

### 2025-03-05 tests_r/test_4_series_DCCA.R

@@ -1,0 +1,36 @@
+require(combinat)
+require("DCCA")
+
+# define time windows scales
+tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
+        362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570);
+print("Number of time windows scales:")
+print(length(tws))
+
+# load data
+data <- read.csv("https://tinyurl.com/zbtestdata01")
+dcca_of <- combn(4, 2)
+
+
+
+result_DCCA_PKG <- list()
+startTime <- Sys.time()
+
+
+for(i in 1:dim(dcca_of)[2]){
+ 
+  
+  s1 <- data[[ dcca_of[,i ][1]]]
+  s2 <- data[[ dcca_of[,i ][2]]]
+  print(c(dcca_of[,i ][1], dcca_of[,i ][2]))
+
+  temp = rhodcca(s1, s2, m = tws, nu = 0, overlap = TRUE)
+  result_DCCA_PKG <- append(result_DCCA_PKG, temp, after = length(result_DCCA_PKG))
+}
+
+endTime <- Sys.time()
+
+# print recorded time
+print("elapsed time:")
+elapsed_DCCA_PKG <- endTime - startTime
+print(elapsed_DCCA_PKG)

### 2025-03-05 tests_r/test_4_series_DFA.R

@@ -1,0 +1,38 @@
+require(combinat)
+require("DFA")
+
+# define time windows scales
+tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
+        362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570);
+print("Number of 
+      
+      time windows scales:")
+print(length(tws))
+
+# load data
+data <- read.csv("https://tinyurl.com/zbtestdata01")
+dcca_of <- combn(4, 2)
+
+
+
+result_DFA_PKG <- list()
+startTime <- Sys.time()
+
+
+for(i in 1:dim(dcca_of)[2]){
+ 
+  
+  s1 <- data[[ dcca_of[,i ][1]]]
+  s2 <- data[[ dcca_of[,i ][2]]]
+  print(c(dcca_of[,i ][1], dcca_of[,i ][2]))
+
+  temp = rhoDCCA(s1, s2, scale ="F", box_size= tws)
+  result_DFA_PKG <- append(result_DFA_PKG, temp , after = length(result_DFA_PKG))
+}
+
+endTime <- Sys.time()
+
+# print recorded time
+print("elapsed time:")
+elapsed_DFA_PKG <- endTime - startTime
+print(elapsed_DFA_PKG)

### 2025-03-06 tests_r/export_bibtex.R
packages citation

@@ -1,0 +1,11 @@
+# Load the knitr package
+library(knitr)
+
+# Specify the packages you want to include in the BibTeX file
+packages <- c("DCCA", "DFA")
+
+# Generate a BibTeX file for the specified packages
+write_bib(x = packages, file = "selected_packages.bib")
+
+# Print a message indicating the file has been created
+cat("BibTeX file 'selected_packages.bib' has been created.\n")

### 2025-03-07 paper_v02/article.tex
comparisson table

--- 2025-03-02 paper_v02/article.tex
@@
 
+The algorithm was tested with data and results from the \cite{RIBEIRO2025107419} paper, the original data came from the \textit{Physionet} open-access website, available at this link:
+
+\begin{center}
+	\url{https://physionet.org/content/eegmmidb/1.0.0/}
+\end{center} 
+This comprehensive dataset of EEG (electroencephalogram) data, with 108 subjects, executing 12 experiments each, with the $\rho_{DCCA}$ and $DMC_{x}^{2}$ calculated for all experiments, using signals from four electrodes.
+The results were compared with the original ones, and the algorithms where validated.
+
+Regarding performance, the pure \proglang{Python} implementation, the \proglang{Zig} implementation and the optimized \proglang{Zig} where compared with 3 oder packages that calculates the $\rho_{DCCA}$.
+Researching the \proglang{R} \textbf{CRAN} package repository, the \pkg{DCCA} \pkg{DFA} packages where found, and, from the \proglang{Python} \textbf{pypi}, the \pkg{DCCA} package was used.
+
+The \pkg{DFA}~\citep{R-DFA} is a pure \proglang{R} implementation.
+The \pkg{DCCA}~\citep{R-DCCA} is a \proglang{R} package implemented in \proglang{Java}.
+A package also named \pkg{DCCA}\citep{alchieri2024lateralization} is a \proglang{Python} implementation with massive use of \pkg{Numpy} broadcasting capabilities and memory access with strides to avoid loops.
+
+\begin{table}[h]
+  \centering
+  \caption{Elapsed Time Comparison}
+  \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
+    \hline
+    Package & Implementation & Elapsed Time(s) & Percentage Increase \\
+    \hline
+    Zebende & Zig (opt) & 27.52  & - \\
+    Zebende & Zig (basic) & 42.03 & 52.7\% \\
+    DCCA & Python & 370.54 & 1239.4\% \\
+    DCCA & R-Java & 1987.67 & 7202.4\% \\
+    Zebende & Pure Python & 2415.94 & 8803.4\% \\
+    DFA & R & 20173.13 & 73,011.4\% \\
+    \hline
+  \end{tabular}
+\end{table}
+
+Regarding performance
+
 %% -- Summary/conclusions/discussion -------------------------------------------

### 2025-03-07 paper_v02/article.tex
time comparing figure

--- 2025-03-02 paper_v02/article.tex
@@
 
-%% -- Summary/conclusions/discussion -------------------------------------------
+The algorithm was tested with data and results from the \cite{RIBEIRO2025107419} paper.
+Details regarding data preparations can be found in the aforementioned paper.
+A comprehensive dataset of EEG (electroencephalogram) data, with 108 subjects, executing 12 experiments each, with the $\rho_{DCCA}$ and $DMC_{x}^{2}$ calculated for all experiments, using signals from four electrodes with 15742 points.
+The results were compared with the original ones, and the algorithms where validated.
+
+The original data came from the \textit{Physionet} open-access website, available at this link:
+
+\begin{center}
+	\url{https://physionet.org/content/eegmmidb/1.0.0/}
+\end{center}
+
+Regarding performance, the pure \proglang{Python} implementation, the \proglang{Zig} implementation and the optimized \proglang{Zig} where compared with 3 oder packages that calculates the $\rho_{DCCA}$.
+Researching the \proglang{R} \textbf{CRAN} repository, the \pkg{DFA}, a pure \proglang{R} implementation, and  the \pkg{DCCA}~\citep{R-DCCA},  a \proglang{R} package implemented in \proglang{Java} where found.
+From the \proglang{Python} \textbf{pypi} repository, a package also named \pkg{DCCA}\citep{alchieri2024lateralization}, a \proglang{Python} implementation with massive use of \pkg{Numpy} broadcasting capabilities and memory access with strides to avoid loops, was used.
+All those packages have $\rho_{DCCA}$ functions.
+Since the $dmc_{x}^{2}$ is a more recent concept, no other package was found that calculates it besides the \pkg{Zebende} package.
+The $\rho_{DCCA}$ is a fundamental step to calculate the $DMC_{x}^{2}$ and contains the most computational expensive parts.
+
+With the idea to broadly compare the performance of the \pkg{Zebende} package with the other packages, aiming an overall order of magnitude of the packages compared speed, using the same computer and the same time window scales, the performance comparison was made using data from one of the experiments described in the \cite{RIBEIRO2025107419} paper, available at the link below:
+
+\begin{center}
+  	\url{https://tinyurl.com/zbtestdata01}
+\end{center}
+
+\begin{table}[h]
+  \label{tab:elapsed_time}
+  \centering
+  \caption{Elapsed Time Comparison}
+  \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
+    \hline
+    Package & Implementation & Elapsed Time(s) & Percentage Increase \\
+    \hline
+    Zebende & Zig (opt) & 27.52  & - \\
+    Zebende & Zig (basic) & 42.03 & 52.7\% \\
+    DCCA & Python & 370.54 & 1239.4\% \\
+    DCCA & R-Java & 1987.67 & 7202.4\% \\
+    Zebende & Python & 2415.94 & 8803.4\% \\
+    DFA & R & 20173.13 & 73,011.4\% \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics{
+    figs/elapsed_tws.png}
+  \caption{\label{fig:elapsed_tws}Elapsed Time Comparison for Different Implementations}
+\end{figure}
 

### 2025-03-10 paper_v02/article.tex
time comparing figure
conclusion

@@
 All those packages have $\rho_{DCCA}$ functions.
-Since the $dmc_{x}^{2}$ is a more recent concept, no other package was found that calculates it besides the \pkg{Zebende} package.
+Hereafter the \pkg{DCCA} \proglang{R-java} package will be referred as \pkg{DCCA-R-Java} and the \pkg{DCCA} \proglang{Python} package as \pkg{DCCA-Python} for better distinction.
+
+The $dmc_{x}^{2}$ is a more recent coefficient, and no other package was found that calculates it besides the \pkg{Zebende} package.
+That alone is a big contribution of this package.
 The $\rho_{DCCA}$ is a fundamental step to calculate the $DMC_{x}^{2}$ and contains the most computational expensive parts.
-
-With the idea to broadly compare the performance of the \pkg{Zebende} package with the other packages, aiming an overall order of magnitude of the packages compared speed, using the same computer and the same time window scales, the performance comparison was made using data from one of the experiments described in the \cite{RIBEIRO2025107419} paper, available at the link below:
+In the $DMC_{x}^{2}$ calculation, The $\rho_{DCCA}$ is calculated for, at least, a combination of 3 series two by two.
+With the idea to broadly compare the performance of the \pkg{Zebende} package with the others, aiming an overall order of magnitude of the packages compared speed, using the same computer and the same time window scales, the performance comparison was made using data from one of the experiments described in the \cite{RIBEIRO2025107419} paper, available at the link below:
 
@@
 
+None of the comparison packages can calculate multiple series comparison in the same input, so the $\rho_{DCCA}$ methods where wrapped around a loop that alternate the series used by the function, using a four two by two combination of the data indexes.
+The \pkg{DCCA-Python} package also didn't accepts multiple inputs for the time scale, so another loop was added in that test.
+The results are presented in the Table \ref{tab:time}.
+
 \begin{table}[h]
-  \label{tab:elapsed_time}
+  \label{tab:time}
   \centering
@@
 
+The \pkg{Zebende} package, using the optimized \proglang{Zig} implementation, was the fastest, with a 27.52 seconds elapsed time.
+The basic \proglang{Zig} implementation was the second fastest, with 42.03 seconds.
+The \pkg{DCCA-Python} package was the third fastest, with 370.54 seconds.
+The \pkg{DCCA-R-Java} package was the fourth fastest, with 1987.67 seconds.
+The \pkg{Zebende} package, using the \proglang{Python} implementation, was the fifth fastest, with 2415.94 seconds.
+The \pkg{DFA-R} package was the slowest, with 20173.13 seconds.
+
+The \proglang{Zig} basic and optimized versions occupying the first two positions in the ranking is not unexpected, hence the \proglang{Zig} language is more performative than the other ones used in the comparison.
+The \pkg{DCCA-Python} package in the third position is a surprise, showing that a well implemented \proglang{Python} package can overperform a \proglang{Java} implementation.
+
+The $52.7\%$ increase in performance from the basic to the optimized \proglang{Zig} implementation is a good result, showcasing the effectiveness of the optimizations.
+Another test was conducted, wth the same data and settings as the previous one, but getting the elapsed time inside every time scale.
+These results are presented in Fig.~\ref{fig:elapsed_tws}.
+
 \begin{figure}[h!]
@@
 
+Figure \ref{fig:elapsed_tws} shows the elapsed computation time difference from the Basic to the optimized version of the algorithm increases as the time scale increases.
+If the test was conducted with a bigger time series, and with larger time scales the difference would be even more significant.
+
 \section{Summary and discussion} \label{sec:summary}
+
+The \pkg{Zebende} package is the first tool that calculates the $DMC_{x}^{2}$ coefficient for any number of time series.
+The package also calculates the $\rho_{DCCA}$, a fundamental step for the $DMC_{x}^{2}$ calculation.
+The package was tested with data from the \cite{RIBEIRO2025107419} paper, and the results were validated.
+The package was also compared with other packages that calculate the $\rho_{DCCA}$, and the results showed that the \pkg{Zebende} package is the fastest.
+The package is available at the \textbf{pypi}.
 

### 2025-03-10 tests_r/test r_DFA_PKG.R

@@ -1,8 +1,11 @@
 require("DFA")
 
 # define time windows scales
-tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
-        362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570)
+tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
+         38, 45, 52, 60, 69, 79, 91, 104, 119,
+         135, 154, 174, 198, 223, 252, 285, 321,
+         362, 407, 457, 513, 575, 645, 723, 809,
+         905, 1011, 1130, 1261, 1407, 1570)
 
 print(length(tws))
 
@@ -14,12 +17,12 @@ f3 <- data[, c("F3")]
 f6 <- data[, c("F6")]
 
 print("Start")
-startTime <- Sys.time() 
+start_time <- Sys.time()
 
-result_DFA_PKG =  rhoDCCA(f3, f6, scale ="F", box_size= tws)
+rhoDCCA(f3, f6, scale = "F", box_size = tws)
 
-endTime <- Sys.time()
+end_time <- Sys.time()
 
 # print recorded time
 print("DFA_PKG result: ")
-print(endTime - startTime)
+print(end_time - start_time)

### 2025-03-10 tests_r/test_4_series_DCCA.R

@@ -1,36 +1,27 @@
 require(combinat)
 require("DCCA")
-
 # define time windows scales
-tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
-        362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570);
+tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
+         38, 45, 52, 60, 69, 79, 91, 104, 119,
+         135, 154, 174, 198, 223, 252, 285, 321,
+         362, 407, 457, 513, 575, 645, 723, 809,
+         905, 1011, 1130, 1261, 1407, 1570)
 print("Number of time windows scales:")
 print(length(tws))
-
-# load data
-data <- read.csv("https://tinyurl.com/zbtestdata01")
+data <- read.csv("https://tinyurl.com/zbtestdata01")# load data
 dcca_of <- combn(4, 2)
-
-
-
 result_DCCA_PKG <- list()
-startTime <- Sys.time()
-
-
+start_time <- Sys.time()
 for(i in 1:dim(dcca_of)[2]){
- 
-  
-  s1 <- data[[ dcca_of[,i ][1]]]
-  s2 <- data[[ dcca_of[,i ][2]]]
-  print(c(dcca_of[,i ][1], dcca_of[,i ][2]))
+  s1 <- data[[ dcca_of[, i][1]]]
+  s2 <- data[[ dcca_of[, i][2]]]
+  print(c(dcca_of[, i][1], dcca_of[, i][2]))
 
-  temp = rhodcca(s1, s2, m = tws, nu = 0, overlap = TRUE)
-  result_DCCA_PKG <- append(result_DCCA_PKG, temp, after = length(result_DCCA_PKG))
+  temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE)
+  result_DCCA_PKG <- append(result_DCCA_PKG, temp, 
+                            after = length(result_DCCA_PKG))
 }
-
-endTime <- Sys.time()
-
-# print recorded time
+end_time <- Sys.time()
 print("elapsed time:")
-elapsed_DCCA_PKG <- endTime - startTime
+elapsed_DCCA_PKG <- end_time - start_time
 print(elapsed_DCCA_PKG)

### 2025-03-10 tests_r/test_4_series_DFA.R

@@ -1,38 +1,28 @@
 require(combinat)
 require("DFA")
-
 # define time windows scales
-tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
-        362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570);
-print("Number of 
-      
-      time windows scales:")
+tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
+         38, 45, 52, 60, 69, 79, 91, 104, 119,
+         135, 154, 174, 198, 223, 252, 285, 321,
+         362, 407, 457, 513, 575, 645, 723, 809,
+         905, 1011, 1130, 1261, 1407, 1570)
+print("Number of time windows scales:")
 print(length(tws))
-
-# load data
-data <- read.csv("https://tinyurl.com/zbtestdata01")
+data <- read.csv("https://tinyurl.com/zbtestdata01") # load data
 dcca_of <- combn(4, 2)
-
-
-
 result_DFA_PKG <- list()
-startTime <- Sys.time()
+start_time <- Sys.time()
+for (i in 1:dim(dcca_of)[2]){
+  s1 <- data[[ dcca_of[, i][1]]]
+  s2 <- data[[ dcca_of[, i][2]]]
+  print(c(dcca_of[, i][1], dcca_of[, i][2]))
 
-
-for(i in 1:dim(dcca_of)[2]){
- 
-  
-  s1 <- data[[ dcca_of[,i ][1]]]
-  s2 <- data[[ dcca_of[,i ][2]]]
-  print(c(dcca_of[,i ][1], dcca_of[,i ][2]))
-
-  temp = rhoDCCA(s1, s2, scale ="F", box_size= tws)
-  result_DFA_PKG <- append(result_DFA_PKG, temp , after = length(result_DFA_PKG))
+  temp <- rhoDCCA(s1, s2, scale = "F", box_size = tws)
+  result_DFA_PKG <- append(result_DFA_PKG, temp,
+                           after = length(result_DFA_PKG))
 }
-
-endTime <- Sys.time()
-
+end_time <- Sys.time()
 # print recorded time
 print("elapsed time:")
-elapsed_DFA_PKG <- endTime - startTime
+elapsed_DFA_PKG <- end_time - start_time
 print(elapsed_DFA_PKG)

### 2025-03-10 paper_v02/article.tex
first draft

@@
 
-% rho DCCA applications
+\color{red}Inserir parágrafo com citações de artigos que aplicam o $\rho_{DCCA}$\color{black}
 
@@
 
-% DMCx2 applications
+The method has been applied to study meteorological data~\citep{Brito2019}, have been tested statistically~\citep{filho2021-2}, used with the implementation of sliding windows~\citep{guedes2021-2}, to measure the contagion effect on stock market indexes~\citep{guedes2022d}, to produce statistical analysis between stock market behavior~\citep{zebende2022s} or analyze electroencephalogram (EEG) signals \citep{RIBEIRO2025107419}.
 
@@
 Details regarding data preparations can be found in the aforementioned paper.
-A comprehensive dataset of EEG (electroencephalogram) data, with 108 subjects, executing 12 experiments each, with the $\rho_{DCCA}$ and $DMC_{x}^{2}$ calculated for all experiments, using signals from four electrodes with 15742 points.
+A comprehensive dataset of EEG, data, with 108 subjects, executing 12 experiments each, with the $\rho_{DCCA}$ and $DMC_{x}^{2}$ calculated for all experiments, using signals from four electrodes with 15742 points.
 The results were compared with the original ones, and the algorithms where validated.

### 2025-03-24 paper_v02/article.tex
updating tests

@@
 
-\color{red}Inserir parágrafo com citações de artigos que aplicam o $\rho_{DCCA}$\color{black}
+The method was statistically validated~\citep{PhysRevE.84.066118}, applied to climate data~\cite{vassolerZebende2012}, financial data\citep{Wang2013, Guedes2017, Ferreira2018}, electroencephalogram recorded data \citep{Oliveira2023}, among others.
+Statistical criteria for evaluating the relevance of the results were also developed~\cite{Guedes2018,Guedes2018a}.
 
@@
   \centering
-  \caption{Elapsed Time Comparison}
+  \caption{Elapsed Time Comparison with four series}
   \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
@@
     figs/elapsed_tws.png}
-  \caption{\label{fig:elapsed_tws}Elapsed Time Comparison for Different Implementations}
+  \caption{\label{fig:elapsed_tws}Elapsed Time Comparison for Different Implementations (four series)}
 \end{figure}

### 2025-03-24 tests_r/test_4_series_DCCA.R

@@ -1,28 +1,0 @@
-require(combinat)
-require("DCCA")
-# define time windows scales
-tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
-         38, 45, 52, 60, 69, 79, 91, 104, 119,
-         135, 154, 174, 198, 223, 252, 285, 321,
-         362, 407, 457, 513, 575, 645, 723, 809,
-         905, 1011, 1130, 1261, 1407, 1570)
-print("Number of time windows scales:")
-print(length(tws))
-series_count <- 6
-data <- read.csv("https://tinyurl.com/zbtestdata64")# load data
-dcca_of <- combn(series_count, 2)
-result_DCCA_PKG <- list()
-start_time <- Sys.time()
-for(i in 1:dim(dcca_of)[2]){
-  s1 <- data[[ dcca_of[, i][1]]]
-  s2 <- data[[ dcca_of[, i][2]]]
-  print(c(dcca_of[, i][1], dcca_of[, i][2]))
-
-  temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE)
-  result_DCCA_PKG <- append(result_DCCA_PKG, temp, 
-                            after = length(result_DCCA_PKG))
-}
-end_time <- Sys.time()
-print("elapsed time:")
-elapsed_DCCA_PKG <- end_time - start_time
-print(elapsed_DCCA_PKG)

### 2025-03-24 tests_r/test_4_series_DFA.R

@@ -1,28 +1,0 @@
-require(combinat)
-require("DFA")
-# define time windows scales
-tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
-         38, 45, 52, 60, 69, 79, 91, 104, 119,
-         135, 154, 174, 198, 223, 252, 285, 321,
-         362, 407, 457, 513, 575, 645, 723, 809,
-         905, 1011, 1130, 1261, 1407, 1570)
-print("Number of time windows scales:")
-print(length(tws))
-data <- read.csv("https://tinyurl.com/zbtestdata01") # load data
-dcca_of <- combn(4, 2)
-result_DFA_PKG <- list()
-start_time <- Sys.time()
-for (i in 1:dim(dcca_of)[2]){
-  s1 <- data[[ dcca_of[, i][1]]]
-  s2 <- data[[ dcca_of[, i][2]]]
-  print(c(dcca_of[, i][1], dcca_of[, i][2]))
-
-  temp <- rhoDCCA(s1, s2, scale = "F", box_size = tws)
-  result_DFA_PKG <- append(result_DFA_PKG, temp,
-                           after = length(result_DFA_PKG))
-}
-end_time <- Sys.time()
-# print recorded time
-print("elapsed time:")
-elapsed_DFA_PKG <- end_time - start_time
-print(elapsed_DFA_PKG)

### 2025-03-24 tests_r/test_n_series_DCCA.R

@@ -1,0 +1,26 @@
+require(combinat)
+require("DCCA")
+# define time windows scales
+tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
+         38, 45, 52, 60, 69, 79, 91, 104, 119,
+         135, 154, 174, 198, 223, 252, 285, 321,
+         362, 407, 457, 513, 575, 645, 723, 809,
+         905, 1011, 1130, 1261, 1407, 1570)
+cat("Number of time windows scales: ", length(tws)
+series_count <- 12 # select the number of series to run 
+data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
+data <- data[, 1:series_count] # filter columns
+start_time <- Sys.time() # start time stamp
+dcca_of <- combn(c(1:series_count), 2)
+cat("Number of series:", length(data)
+result_DCCA_PKG <- list()
+for(i in 1:dim(dcca_of)[2]){
+  s1 <- data[[ dcca_of[, i][1]]] # selecting first seires
+  s2 <- data[[ dcca_of[, i][2]]] # selecting second seires
+  temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE) # Pdcca calculations
+  result_DCCA_PKG <- append(result_DCCA_PKG, temp,
+                            after = length(result_DCCA_PKG)) # collecting results
+}
+end_time <- Sys.time() #end time stamp
+elapsed_DCCA_PKG <- end_time - start_time #elapsed time
+cat("elapsed time: ", elapsed_DCCA_PKG)

### 2025-03-24 tests_r/test_n_series_DFA.R

@@ -1,0 +1,24 @@
+require(combinat)
+require("DFA")
+# define time windows scales
+tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
+         38, 45, 52, 60, 69, 79, 91, 104, 119,
+         135, 154, 174, 198, 223, 252, 285, 321,
+         362, 407, 457, 513, 575, 645, 723, 809,
+         905, 1011, 1130, 1261, 1407, 1570)
+cat("Number of time windows scales: ",length(tws)
+series_count <- 12 # select the number of series to run 
+data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
+data <- data[, 1:series_count] # filter columns
+start_time <- Sys.time() # time stamp
+dcca_of <- combn(c(1:series_count), 2)
+print("Number of series:"); print(length(data))
+for (i in 1:dim(dcca_of)[2]){
+  s1 <- data[[ dcca_of[, i][1]]] # selecting first seires
+  s2 <- data[[ dcca_of[, i][2]]] # selecting second seires
+
+  rhoDCCA(s1, s2, scale = "F", box_size = tws) # Pdcca calculations
+}
+end_time <- Sys.time() #end time stamp
+elapsed_DFA_PKG <- end_time - start_time #elapsed time
+cat("elapsed time: ", elapsed_DFA_PKG)

### 2025-03-24 tests_r/test_n_series_DCCA.R
calc

@@ -6,13 +6,13 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          135, 154, 174, 198, 223, 252, 285, 321,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
-cat("Number of time windows scales: ", length(tws)
+cat("Number of time windows scales: ", length(tws), "\n")
 series_count <- 12 # select the number of series to run 
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # start time stamp
 dcca_of <- combn(c(1:series_count), 2)
-cat("Number of series:", length(data)
+cat("Number of series:", length(data), "\n")
 result_DCCA_PKG <- list()
 for(i in 1:dim(dcca_of)[2]){
   s1 <- data[[ dcca_of[, i][1]]] # selecting first seires
@@ -23,4 +23,4 @@ for(i in 1:dim(dcca_of)[2]){
 }
 end_time <- Sys.time() #end time stamp
 elapsed_DCCA_PKG <- end_time - start_time #elapsed time
-cat("elapsed time: ", elapsed_DCCA_PKG)
+cat("elapsed time: ", elapsed_DCCA_PKG, "\n")

### 2025-03-24 tests_r/test_n_series_DFA.R

@@ -6,13 +6,13 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          135, 154, 174, 198, 223, 252, 285, 321,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
-cat("Number of time windows scales: ",length(tws)
+cat("Number of time windows scales: ",length(tws), "\n")
 series_count <- 12 # select the number of series to run 
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # time stamp
 dcca_of <- combn(c(1:series_count), 2)
-print("Number of series:"); print(length(data))
+cat("Number of series: ", length(data), "\n")
 for (i in 1:dim(dcca_of)[2]){
   s1 <- data[[ dcca_of[, i][1]]] # selecting first seires
   s2 <- data[[ dcca_of[, i][2]]] # selecting second seires
@@ -21,4 +21,4 @@ for (i in 1:dim(dcca_of)[2]){
 }
 end_time <- Sys.time() #end time stamp
 elapsed_DFA_PKG <- end_time - start_time #elapsed time
-cat("elapsed time: ", elapsed_DFA_PKG)
+cat("elapsed time: ", elapsed_DFA_PKG, "\n")

### 2025-03-24 tests_r/test r_DCCA_PKG.R
update tests

@@ -1,30 +1,0 @@
-require("DCCA")
-
-# define time windows scales
-tws = c(4, 5 , 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
-         362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570);
-
-print(length(tws));
-
-# load data
-data <- read.csv("https://tinyurl.com/zbtestdata01")
-
-# extract series
-f3 <- data[, 1];
-f6 <- data[, c("F6")];
-print(f3)
-print("Start")
-startTime <- Sys.time() 
-
-result_DCCA_PKG =  rhodcca(f3, f6, m = tws, nu = 0, overlap = TRUE)
-
-endTime <- Sys.time()
-
-# print recorded time 
-print(endTime - startTime)
-
-result_DCCA_PKG_pdcca = result_DCCA_PKG[["rhodcca"]]
-
-
-
-

### 2025-03-24 tests_r/test r_DFA_PKG.R

@@ -1,28 +1,0 @@
-require("DFA")
-
-# define time windows scales
-tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
-         38, 45, 52, 60, 69, 79, 91, 104, 119,
-         135, 154, 174, 198, 223, 252, 285, 321,
-         362, 407, 457, 513, 575, 645, 723, 809,
-         905, 1011, 1130, 1261, 1407, 1570)
-
-print(length(tws))
-
-# read data
-data <- read.csv("https://tinyurl.com/zbtestdata01")
-
-# extract series
-f3 <- data[, c("F3")]
-f6 <- data[, c("F6")]
-
-print("Start")
-start_time <- Sys.time()
-
-rhoDCCA(f3, f6, scale = "F", box_size = tws)
-
-end_time <- Sys.time()
-
-# print recorded time
-print("DFA_PKG result: ")
-print(end_time - start_time)

### 2025-03-24 tests_r/test_n_series_DCCA.R

@@ -7,7 +7,7 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
 cat("Number of time windows scales: ", length(tws), "\n")
-series_count <- 12 # select the number of series to run 
+series_count <- 4 # select the number of series to run 
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # start time stamp
@@ -15,8 +15,8 @@ dcca_of <- combn(c(1:series_count), 2)
 cat("Number of series:", length(data), "\n")
 result_DCCA_PKG <- list()
 for(i in 1:dim(dcca_of)[2]){
-  s1 <- data[[ dcca_of[, i][1]]] # selecting first seires
-  s2 <- data[[ dcca_of[, i][2]]] # selecting second seires
+  s1 <- data[[ dcca_of[, i][1]]] # selecting first series
+  s2 <- data[[ dcca_of[, i][2]]] # selecting second series
   temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE) # Pdcca calculations
   result_DCCA_PKG <- append(result_DCCA_PKG, temp,
                             after = length(result_DCCA_PKG)) # collecting results

### 2025-03-24 tests_r/test_n_series_DFA.R

@@ -7,15 +7,15 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
 cat("Number of time windows scales: ",length(tws), "\n")
-series_count <- 12 # select the number of series to run 
+series_count <- 4 # select the number of series to run
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # time stamp
 dcca_of <- combn(c(1:series_count), 2)
 cat("Number of series: ", length(data), "\n")
 for (i in 1:dim(dcca_of)[2]){
-  s1 <- data[[ dcca_of[, i][1]]] # selecting first seires
-  s2 <- data[[ dcca_of[, i][2]]] # selecting second seires
+  s1 <- data[[ dcca_of[, i][1]]] # selecting first series
+  s2 <- data[[ dcca_of[, i][2]]] # selecting second series
 
   rhoDCCA(s1, s2, scale = "F", box_size = tws) # Pdcca calculations
 }

### 2025-03-24 tests_r/test_n_series_DCCA.R
update tests

@@ -14,9 +14,9 @@ start_time <- Sys.time() # start time stamp
 dcca_of <- combn(c(1:series_count), 2)
 cat("Number of series:", length(data), "\n")
 result_DCCA_PKG <- list()
-for(i in 1:dim(dcca_of)[2]){
-  s1 <- data[[ dcca_of[, i][1]]] # selecting first series
-  s2 <- data[[ dcca_of[, i][2]]] # selecting second series
+for(i in seq_len(dim(dcca_of)[2])){
+  s1 <- data[[dcca_of[, i][1]]] # selecting first series
+  s2 <- data[[dcca_of[, i][2]]] # selecting second series
   temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE) # Pdcca calculations
   result_DCCA_PKG <- append(result_DCCA_PKG, temp,
                             after = length(result_DCCA_PKG)) # collecting results

### 2025-03-24 tests_r/test_n_series_DFA.R

@@ -6,16 +6,16 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          135, 154, 174, 198, 223, 252, 285, 321,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
-cat("Number of time windows scales: ",length(tws), "\n")
+cat("Number of time windows scales: ", length(tws), "\n")
 series_count <- 4 # select the number of series to run
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # time stamp
 dcca_of <- combn(c(1:series_count), 2)
 cat("Number of series: ", length(data), "\n")
-for (i in 1:dim(dcca_of)[2]){
-  s1 <- data[[ dcca_of[, i][1]]] # selecting first series
-  s2 <- data[[ dcca_of[, i][2]]] # selecting second series
+for (i in seq_len(dim(dcca_of)[2])){
+  s1 <- data[[dcca_of[, i][1]]] # selecting first series
+  s2 <- data[[dcca_of[, i][2]]] # selecting second series
 
   rhoDCCA(s1, s2, scale = "F", box_size = tws) # Pdcca calculations
 }

### 2025-03-24 tests_r/test_n_series_DCCA.R
update tests

@@ -7,20 +7,20 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
 cat("Number of time windows scales: ", length(tws), "\n")
-series_count <- 4 # select the number of series to run 
+series_count <- 4 # select the number of series to run
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # start time stamp
 dcca_of <- combn(c(1:series_count), 2)
 cat("Number of series:", length(data), "\n")
-result_DCCA_PKG <- list()
+result_dcca_pkg <- list()
 for(i in seq_len(dim(dcca_of)[2])){
   s1 <- data[[dcca_of[, i][1]]] # selecting first series
   s2 <- data[[dcca_of[, i][2]]] # selecting second series
   temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE) # Pdcca calculations
-  result_DCCA_PKG <- append(result_DCCA_PKG, temp,
-                            after = length(result_DCCA_PKG)) # collecting results
+  result_dcca_pkg <- append(result_DCCA_PKG, temp,  # collecting results
+                            after = length(result_DCCA_PKG))
 }
 end_time <- Sys.time() #end time stamp
-elapsed_DCCA_PKG <- end_time - start_time #elapsed time
-cat("elapsed time: ", elapsed_DCCA_PKG, "\n")
+elapsed_dcca_pkg <- end_time - start_time #elapsed time
+cat("elapsed time: ", elapsed_dcca_pkg, "\n")

### 2025-03-24 tests_r/test_n_series_DFA.R

@@ -16,9 +16,8 @@ cat("Number of series: ", length(data), "\n")
 for (i in seq_len(dim(dcca_of)[2])){
   s1 <- data[[dcca_of[, i][1]]] # selecting first series
   s2 <- data[[dcca_of[, i][2]]] # selecting second series
-
   rhoDCCA(s1, s2, scale = "F", box_size = tws) # Pdcca calculations
 }
 end_time <- Sys.time() #end time stamp
-elapsed_DFA_PKG <- end_time - start_time #elapsed time
-cat("elapsed time: ", elapsed_DFA_PKG, "\n")
+elapsed_dfa_pkg <- end_time - start_time #elapsed time
+cat("elapsed time: ", elapsed_dfa_pkg, "\n")

### 2025-03-24 tests_r/test_n_series_DCCA.R
update tests

@@ -17,6 +17,7 @@ result_dcca_pkg <- list()
 for(i in seq_len(dim(dcca_of)[2])){
   s1 <- data[[dcca_of[, i][1]]] # selecting first series
   s2 <- data[[dcca_of[, i][2]]] # selecting second series
+  cat("Series: ", dcca_of[, i][1], " and ", dcca_of[, i][2], "\n")
   temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE) # Pdcca calculations
   result_dcca_pkg <- append(result_DCCA_PKG, temp,  # collecting results
                             after = length(result_DCCA_PKG))

### 2025-03-24 tests_r/test_n_series_DFA.R

@@ -16,6 +16,7 @@ cat("Number of series: ", length(data), "\n")
 for (i in seq_len(dim(dcca_of)[2])){
   s1 <- data[[dcca_of[, i][1]]] # selecting first series
   s2 <- data[[dcca_of[, i][2]]] # selecting second series
+  cat("Series: ", dcca_of[, i][1], " and ", dcca_of[, i][2], "\n")
   rhoDCCA(s1, s2, scale = "F", box_size = tws) # Pdcca calculations
 }
 end_time <- Sys.time() #end time stamp

### 2025-03-25 tests_r/test_n_series_DCCA.R
update tests

@@ -1,5 +1,6 @@
 require(combinat)
 require("DCCA")
+series_count <- 4 # select the number of series to run
 # define time windows scales
 tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          38, 45, 52, 60, 69, 79, 91, 104, 119,
@@ -7,7 +8,6 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
 cat("Number of time windows scales: ", length(tws), "\n")
-series_count <- 4 # select the number of series to run
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # start time stamp
@@ -19,8 +19,8 @@ for(i in seq_len(dim(dcca_of)[2])){
   s2 <- data[[dcca_of[, i][2]]] # selecting second series
   cat("Series: ", dcca_of[, i][1], " and ", dcca_of[, i][2], "\n")
   temp <- rhodcca(s1, s2, m = tws, nu = 1, overlap = TRUE) # Pdcca calculations
-  result_dcca_pkg <- append(result_DCCA_PKG, temp,  # collecting results
-                            after = length(result_DCCA_PKG))
+  result_dcca_pkg <- append(result_dcca_pkg, temp,  # collecting results
+                            after = length(result_dcca_pkg))
 }
 end_time <- Sys.time() #end time stamp
 elapsed_dcca_pkg <- end_time - start_time #elapsed time

### 2025-03-25 tests_r/test_n_series_DFA.R

@@ -1,5 +1,6 @@
 require(combinat)
 require("DFA")
+series_count <- 4 # select the number of series to run
 # define time windows scales
 tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          38, 45, 52, 60, 69, 79, 91, 104, 119,
@@ -7,7 +8,6 @@ tws <- c(4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33,
          362, 407, 457, 513, 575, 645, 723, 809,
          905, 1011, 1130, 1261, 1407, 1570)
 cat("Number of time windows scales: ", length(tws), "\n")
-series_count <- 4 # select the number of series to run
 data <- read.csv("https://tinyurl.com/zbtestdata64c")# load data
 data <- data[, 1:series_count] # filter columns
 start_time <- Sys.time() # time stamp

### 2025-04-05 paper_v02/article.tex
update latex table

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$ methods}
+\title{A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$}
 \Plaintitle{A Python/Zig optimized and customizable implementation for the Pdcca and DMCx2 methods}
@@
 This paper presents the \pkg{Zebende} \proglang{Python} package, an implementation of the $DFA$, $DCCA$, $\rho_{DCCCA}$, $DMC_x^2$ and utility functions related to the methods.
-In section \ref{sec:calculations} the steps for calculating the $\rho_{DCCCA}$ and $DMC_x^2$ are presented and discussed.
+In section \ref{sec:calculations} the steps for calculating the $\rho_{DCCA}$ and $DMC_x^2$ are presented and discussed.
 Section \ref{sec:optimization} shows how this library was implemented, the optimization technics and the recommended steps to use the library.
@@
 The \pkg{DCCA-Python} package also didn't accepts multiple inputs for the time scale, so another loop was added in that test.
-The results are presented in the Table \ref{tab:time}.
+The results are presented in the Table \ref{tab:time_4}.
 
   \begin{table}[h]
-  \label{tab:time}
+
     \centering
-  \caption{Elapsed Time Comparison with four series}
+    \caption{Elapsed Time Comparison with four series} \label{tab:time_4}
     \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
@@
       \hline
-    Zebende & Zig (opt) & 27.52  & - \\
-    Zebende & Zig (basic) & 42.03 & 52.7\% \\
-    DCCA & Python & 370.54 & 1239.4\% \\
-    DCCA & R-Java & 1987.67 & 7202.4\% \\
-    Zebende & Python & 2415.94 & 8803.4\% \\
-    DFA & R & 20173.13 & 73,011.4\% \\
+      Zebende & Zig (opt) & 35.88  & - \\
+      Zebende & Zig (basic) & 59.74 & 52.7\% \\
+      DCCA & Python & 447.67 &  1,243.49\% \\
+      DCCA & R-Java & 2,206.3 & 6,126.17\% \\
+      Zebende & Python & 3,076.23 & 8,542.83\% \\
+      DFA & R & 28,818.016 & 80,049.42\% \\
+      \hline
+    \end{tabular}
+  \end{table}
+
+    \begin{table}[h]
+
+      \centering
+      \caption{Elapsed Time Comparison with twelve series} \label{tab:time_12}
+      \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
+        \hline
+        Package & Implementation & Elapsed Time(s) & Percentage Increase \\
+        \hline
+        Zebende & Zig (opt) & 120.55  & - \\
+        Zebende & Zig (basic) & 213.08 & 76.64\% \\
+        Zebende & Python & 3,761.40 & 3,085.47\% \\
+        DCCA & Python & 4,675.17 &  3,871.65\% \\
+        DCCA & R-Java & 4,8624.44 & 40,423.52\% \\
+        DFA & R & 2,060,776.24 & 81,704,499.33\% \\
         \hline

### 2025-04-06 notebooks/zebende_zig_tests.py
zig tests

@@ -1,0 +1,41 @@
+# %%
+import numpy as np 
+import pandas as pd
+import matplotlib.pyplot as plt
+import zebende as zb
+import time
+
+# %%
+series_count = 4
+
+# %%
+url = 'https://tinyurl.com/zbtestdata64c'
+
+# %%
+df = pd.read_csv(url)
+print(df.shape)
+df = df.iloc[:, :series_count]
+df.shape
+
+# %%
+tws =  np.array([4, 5, 7, 9, 11, 13, 16, 20, 23, 28, 33, 38, 45, 52, 60, 69, 79, 91, 104, 119, 135, 154, 174, 198, 223, 252, 285, 321,
+  362, 407, 457, 513, 575, 645, 723, 809, 905, 1011, 1130, 1261, 1407, 1570])
+
+# %%
+df.info()
+
+# %%
+mat = df.to_numpy()
+print(mat)
+
+# %%
+start_time = time.time()
+int_data = zb.integrated_series(mat)
+dfa, dcca, pdcca = zb.p_dcca(mat, tws) 
+end_time = time.time()
+print(f"Time taken: {end_time - start_time} seconds")
+
+# %%
+pdcca
+
+

### 2025-04-06 notebooks/zebende_zig_tests.py
zig tests

@@ -6,7 +6,7 @@ import zebende as zb
 import time
 
 # %%
-series_count = 4
+series_count = 12
 
 # %%
 url = 'https://tinyurl.com/zbtestdata64c'

### 2025-04-06 paper_v02/article.tex
zig tests
latex time tables

@@
 %% - \AND starts a new line, \And does not.
-\author{Fernando Ferraz Ribeiro~\orcidlink{0000-0002-0685-4774}\\Universidade Federal da Bahia
+\author{Fernando Ferraz Ribeiro~\orcidlink{0000-0002-0685-4774}\\Federal University of Bahia
    \And Gilney Figueira Zebende~\orcidlink{0000-0003-2420-9805}\\Universidade Estadual\\de Feira de Santana}
@@
       Zebende & Zig (opt) & 35.88  & - \\
-      Zebende & Zig (basic) & 59.74 & 52.7\% \\
-      DCCA & Python & 447.67 &  1,243.49\% \\
-      DCCA & R-Java & 2,206.3 & 6,126.17\% \\
-      Zebende & Python & 3,076.23 & 8,542.83\% \\
-      DFA & R & 28,818.016 & 80,049.42\% \\
+      Zebende & Zig (basic) & 59.74 & 66.50\% \\
+      DCCA & Python & 447.67 &  1,147.69\% \\
+      DCCA & R-Java & 2,206.2 & 6,048.83\% \\
+      Zebende & Python & 3,076.23 & 8,473.66\% \\
+      DFA & R & 28,116.00 & 78,261.20\% \\
       \hline
@@
         Zebende & Zig (basic) & 213.08 & 76.64\% \\
-        Zebende & Python & 3,761.40 & 3,085.47\% \\
-        DCCA & Python & 4,675.17 &  3,871.65\% \\
-        DCCA & R-Java & 4,8624.44 & 40,423.52\% \\
-        DFA & R & 2,060,776.24 & 81,704,499.33\% \\
+        Zebende & Python & 3,761.40 & 3,020.20\% \\
+        DCCA & Python & 4,675.17 &  3,778.20\% \\
+        DCCA & R-Java & 28,980.00 & 23,939.82\% \\
+        DFA & R & 309,276.00 & 256,454.13\% \\
         \hline
@@
   \includegraphics{
-    figs/elapsed_tws.png}
-  \caption{\label{fig:elapsed_tws}Elapsed Time Comparison for Different Implementations (four series)}
+    figs/elapsed_tws_04.png}
+  \caption{\label{fig:elapsed_tws_04}Elapsed Time Comparison for Different Implementations (four time series)}
 \end{figure}
@@
 If the test was conducted with a bigger time series, and with larger time scales the difference would be even more significant.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics{
+    figs/elapsed_tws_12.png}
+  \caption{\label{fig:elapsed_tws_12}Elapsed Time Comparison for Different Implementations (twelve time series)}
+\end{figure}
 

### 2025-04-06 paper_v02/article.tex
latex time tables

@@
 Another test was conducted, wth the same data and settings as the previous one, but getting the elapsed time inside every time scale.
-These results are presented in Fig.~\ref{fig:elapsed_tws}.
+These results are presented in Fig.~\ref{fig:elapsed_tws_04}.
 
@@
 
-Figure \ref{fig:elapsed_tws} shows the elapsed computation time difference from the Basic to the optimized version of the algorithm increases as the time scale increases.
+Figure \ref{fig:elapsed_tws_04
+} shows the elapsed computation time difference from the Basic to the optimized version of the algorithm increases as the time scale increases.
 If the test was conducted with a bigger time series, and with larger time scales the difference would be even more significant.

### 2025-04-06 paper_v02/article.tex
elapsed time images

@@
 
-Figure \ref{fig:elapsed_tws_04
-} shows the elapsed computation time difference from the Basic to the optimized version of the algorithm increases as the time scale increases.
+Figure \ref{fig:elapsed_tws_04} shows the elapsed computation time difference from the Basic to the optimized version of the algorithm increases as the time scale increases.
 If the test was conducted with a bigger time series, and with larger time scales the difference would be even more significant.

### 2025-04-08 paper_v02/article.tex
elapsed time images
12 series paper description

@@
 The results were compared with the original ones, and the algorithms where validated.
-
-The original data came from the \textit{Physionet} open-access website, available at this link:
+The original data came from the~\cite{schalk2004bci2000} paper, available on the \textit{Physionet} open-access website, that can be accessed in this link:
 
@@
 Regarding performance, the pure \proglang{Python} implementation, the \proglang{Zig} implementation and the optimized \proglang{Zig} where compared with 3 oder packages that calculates the $\rho_{DCCA}$.
-Researching the \proglang{R} \textbf{CRAN} repository, the \pkg{DFA}, a pure \proglang{R} implementation, and  the \pkg{DCCA}~\citep{R-DCCA},  a \proglang{R} package implemented in \proglang{Java} where found.
-From the \proglang{Python} \textbf{pypi} repository, a package also named \pkg{DCCA}\citep{alchieri2024lateralization}, a \proglang{Python} implementation with massive use of \pkg{Numpy} broadcasting capabilities and memory access with strides to avoid loops, was used.
+Researching the \proglang{R} \textbf{CRAN} repository, the \pkg{DFA}~\citep{R-DFA}, a pure \proglang{R} implementation, and  the \pkg{DCCA}~\citep{R-DCCA},  a \proglang{R} package implemented in \proglang{Java} where found.
+From the \proglang{Python} \textbf{pypi} repository, a package also named \pkg{DCCA}\citep{alchieri2024lateralization}, a \proglang{Python} implementation with massive use of \pkg{Numpy} broadcasting capabilities and memory access with strides to avoid loops, was also used.
 All those packages have $\rho_{DCCA}$ functions.
@@
 The $dmc_{x}^{2}$ is a more recent coefficient, and no other package was found that calculates it besides the \pkg{Zebende} package.
-That alone is a big contribution of this package.
+That alone is a huge contribution of this package.
 The $\rho_{DCCA}$ is a fundamental step to calculate the $DMC_{x}^{2}$ and contains the most computational expensive parts.
 In the $DMC_{x}^{2}$ calculation, The $\rho_{DCCA}$ is calculated for, at least, a combination of 3 series two by two.
-With the idea to broadly compare the performance of the \pkg{Zebende} package with the others, aiming an overall order of magnitude of the packages compared speed, using the same computer and the same time window scales, the performance comparison was made using data from one of the experiments described in the \cite{RIBEIRO2025107419} paper, available at the link below:
+With the idea to broadly compare the performance of the \pkg{Zebende} package with the others, aiming an overall order of magnitude of the packages compared speed, using the same computer and the same time window scales, the performance comparison was made using data from one of the experiments described in the~\cite{schalk2004bci2000} paper, with $19920$ lines and $64$ channels.
+For convenience the data was posted in the link below:
 
@@
 The \pkg{DCCA-Python} package also didn't accepts multiple inputs for the time scale, so another loop was added in that test.
-The results are presented in the Table \ref{tab:time_4}.
 
-  \begin{table}[h]
+First all the packages were used to calculate the $\rho_{DCCA}$ of combinations of four time series. The results are presented in the Table \ref{tab:time_4}.
 
+  \begin{table}[h!]
     \centering
@@
       \hline
-      Package & Implementation & Elapsed Time(s) & Percentage Increase \\
+      Package & Implementation & Elapsed Time(s) & Performance Increase \\
       \hline
@@
       DFA & R & 28,116.00 & 78,261.20\% \\
-      \hline
-    \end{tabular}
-  \end{table}
-
-    \begin{table}[h]
-
-      \centering
-      \caption{Elapsed Time Comparison with twelve series} \label{tab:time_12}
-      \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
-        \hline
-        Package & Implementation & Elapsed Time(s) & Percentage Increase \\
-        \hline
-        Zebende & Zig (opt) & 120.55  & - \\
-        Zebende & Zig (basic) & 213.08 & 76.64\% \\
-        Zebende & Python & 3,761.40 & 3,020.20\% \\
-        DCCA & Python & 4,675.17 &  3,778.20\% \\
-        DCCA & R-Java & 28,980.00 & 23,939.82\% \\
-        DFA & R & 309,276.00 & 256,454.13\% \\
       \hline
@@
 
-The $52.7\%$ increase in performance from the basic to the optimized \proglang{Zig} implementation is a good result, showcasing the effectiveness of the optimizations.
-Another test was conducted, wth the same data and settings as the previous one, but getting the elapsed time inside every time scale.
-These results are presented in Fig.~\ref{fig:elapsed_tws_04}.
-
 \begin{figure}[h!] 
@@
 
-Figure \ref{fig:elapsed_tws_04} shows the elapsed computation time difference from the Basic to the optimized version of the algorithm increases as the time scale increases.
+The $66.50\%$ increase in performance from the basic to the optimized \proglang{Zig} implementation is a good result, showcasing the effectiveness of the optimizations.
+Another test was conducted, wth the same data and settings as the previous one, but getting the elapsed time inside every time scale.
+These results are presented in Fig.~\ref{fig:elapsed_tws_04}.
+
+Figure \ref{fig:elapsed_tws_04} shows the elapsed computation time for the Basic and the optimized version of the algorithm.
+The differences between the two implementations increases as the time scale increases.
 If the test was conducted with a bigger time series, and with larger time scales the difference would be even more significant.
+
+The \pkg{Zebende} package implementations where designed to increase performance in scenarios with large number of points in the series, but also with a large amount of time series used in the $DMC_x^2$ calculations.
+To prove that, the same tests were also made using the same data, with the combinations of twelve series, 2 by 2.
+The rasults of that second test are presented in Table~\ref{tab:time_12}.
+
+\begin{table}[h!]
+
+  \centering
+  \caption{Elapsed Time Comparison with twelve series} \label{tab:time_12}
+  \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
+    \hline
+    Package & Implementation & Elapsed Time(s) & Performance Increase \\
+    \hline
+    Zebende & Zig (opt) & 120.55  & - \\
+    Zebende & Zig (basic) & 213.08 & 76.64\% \\
+    Zebende & Python & 3,761.40 & 3,020.20\% \\
+    DCCA & Python & 4,675.17 &  3,778.20\% \\
+    DCCA & R-Java & 28,980.00 & 23,939.82\% \\
+    DFA & R & 309,276.00 & 256,454.13\% \\
+    \hline
+  \end{tabular}
+\end{table}
+
+With that configuration, the results are very simmilar, except that the \pkg{Zebende} pure \proglang{Python} implementation ended in the third position, performing better than the \pkg{DCCA-Python} (4th) and the \pkg{DCCA-R-Java} (5th) packages.
 

### 2025-04-08 paper_v02/article.tex
12 series paper description

@@
 
-With that configuration, the results are very simmilar, except that the \pkg{Zebende} pure \proglang{Python} implementation ended in the third position, performing better than the \pkg{DCCA-Python} (4th) and the \pkg{DCCA-R-Java} (5th) packages.
+With that configuration, the optimized \proglang{Zig} version still the fastest, followed by the basic \proglang{Zig} implementation.
+The \pkg{Zebende} pure \proglang{Python} implementation ended in the third position, performing better than the \pkg{DCCA-Python} (4th) and the \pkg{DCCA-R-Java} (5th) packages.
+The lists ended with the \pkg{DFA} in the 6th position.
 

### 2025-04-08 paper_v02/article.tex
conclusion

@@
 
+  \begin{figure}[h!] 
+    \centering
+    \includegraphics{
+      figs/elapsed_tws_04.png}
+    \caption{\label{fig:elapsed_tws_04}Elapsed Time Comparison for Different Implementations (four time series)}
+  \end{figure}
+
 The \pkg{Zebende} package, using the optimized \proglang{Zig} implementation, was the fastest, with a 27.52 seconds elapsed time.
@@
 
-\begin{figure}[h!] 
-  \centering
-  \includegraphics{
-    figs/elapsed_tws_04.png}
-  \caption{\label{fig:elapsed_tws_04}Elapsed Time Comparison for Different Implementations (four time series)}
-\end{figure}
-
-The $66.50\%$ increase in performance from the basic to the optimized \proglang{Zig} implementation is a good result, showcasing the effectiveness of the optimizations.
+The $66.50\%$ increase in performance in the basic compared to the optimized \proglang{Zig} implementation is a good result, showcasing the effectiveness of the optimizations.
 Another test was conducted, wth the same data and settings as the previous one, but getting the elapsed time inside every time scale.
@@
 To prove that, the same tests were also made using the same data, with the combinations of twelve series, 2 by 2.
-The rasults of that second test are presented in Table~\ref{tab:time_12}.
+The results of that second test are presented in Table~\ref{tab:time_12}.
 
@@
 
-With that configuration, the optimized \proglang{Zig} version still the fastest, followed by the basic \proglang{Zig} implementation.
-The \pkg{Zebende} pure \proglang{Python} implementation ended in the third position, performing better than the \pkg{DCCA-Python} (4th) and the \pkg{DCCA-R-Java} (5th) packages.
-The lists ended with the \pkg{DFA} in the 6th position.
-
 \begin{figure}[h!]
@@
 
+With twelve time series, the optimized \proglang{Zig} version still the fastest, followed by the basic \proglang{Zig} implementation.
+The \pkg{Zebende} pure \proglang{Python} implementation ended in the third position, performing better than the \pkg{DCCA-Python} (4th) and the \pkg{DCCA-R-Java} (5th) packages.
+The lists ended with the \pkg{DFA} in the 6th position.
+
+Comparing the basic with the optimized \proglang{Zig} algorithms using twelve time series the performance increase is $76.64\%$.
+Figure \ref{fig:elapsed_tws_12} shows the comparison between the basic and optimized \proglang{Zig} implementations with twelve time series.
+The results are equivalent to the ones obtained with four time series, reinforcing the capabilities of the optimization technics.
+
+The codes used for testing are available in \url{replace}, as jupyter noteooks, in  and can easily be tested in Google colab environment.
+The basic version is available in \pkg{Zebende} version $0.1.91$ and the optimized implementation in version $0.2.0$ of the package.
+
 \section{Summary and discussion} \label{sec:summary}
@@
 The \pkg{Zebende} package is the first tool that calculates the $DMC_{x}^{2}$ coefficient for any number of time series.
-The package also calculates the $\rho_{DCCA}$, a fundamental step for the $DMC_{x}^{2}$ calculation.
+The package also calculates the $DFA$, $DCCA$ and the $\rho_{DCCA}$, a fundamental step for the $DMC_{x}^{2}$ calculation.
 The package was tested with data from the \cite{RIBEIRO2025107419} paper, and the results were validated.
-The package was also compared with other packages that calculate the $\rho_{DCCA}$, and the results showed that the \pkg{Zebende} package is the fastest.
+The package was also compared with other packages that calculate the $\rho_{DCCA}$, and the results showed that the \pkg{Zebende} package is the fastest, making it suitable for large datasets.
+
+The tests with twelve time scales also shows that the \pkg{Zebene} implementation, designed to obtain better performance working with greater number of time series, deliver great results even with the pure \proglang{Python} non-optimized version.
+This package enable the exploration of the $DMC_x^2$ coefficient in scenarios where it was never tested before.
 The package is available at the \textbf{pypi}.

### 2025-04-08 paper_v02/article.tex
tests page link

@@
 
-The codes used for testing are available in \url{replace}, as jupyter noteooks, in  and can easily be tested in Google colab environment.
 The basic version is available in \pkg{Zebende} version $0.1.91$ and the optimized implementation in version $0.2.0$ of the package.
+The codes used for testing are available, as jupyter noteooks, in the link below, where they can be tested in Google colab environment:
+
+\begin{center}
+  \url{https://255ribeiro.github.io/paper_zebendelib/notebooks/test_nb.html}
+\end{center}
 

### 2025-04-08 paper_v02/article.tex
tests page link
article first version finished

@@
 \author{Fernando Ferraz Ribeiro~\orcidlink{0000-0002-0685-4774}\\Federal University of Bahia
-   \And Gilney Figueira Zebende~\orcidlink{0000-0003-2420-9805}\\Universidade Estadual\\de Feira de Santana}
+   \And Gilney Figueira Zebende~\orcidlink{0000-0003-2420-9805}\\State University\\of Feira de Santana}
 \Plainauthor{Fernando Ferraz Ribeiro, Gilney Figueira Zebende}
@@
 \Abstract{
-  This paper presents tha \pkg{Zebende}, a \proglang{Python} package written in \proglang{Python} and \proglang{Zig}, that calculates the $DFA$, $DCCA$ $\rho_{DCCA}$ and the $DMC_x^2$.
-The package presents an optimized algorithm that significantly improves the calculations speed.
-A comparison with other packages that calculates the $\rho_{DCCA}$.
-The package is also the first to implement the $DMC_x^2$ coefficient in a package and the first algorithm to calculates it for any number of time series.
-}
+  This paper presents \pkg{Zebende}, a \proglang{Python} package written in \proglang{Python} and \proglang{Zig}, that calculates the $DFA$, $DCCA$ $\rho_{DCCA}$ and the $DMC_x^2$, along with auxiliary functions to help the calculations process.
+The package presents an optimized algorithm that significantly improves the calculations speed in large datasets.
+\pkg{Zebende} is the first package that implements the $DMC_x^2$ and can be applied to any number of time series.
+The optimized version of the algorithm is compared to the standard implementation, presenting valuable performance increase.
+A comparison with other \proglang{Python} and \proglang{R} packages that calculates the $\rho_{DCCA}$ is also presented.}
 
@@
   Fernando Ferraz Ribeiro\\
-  Universidade Federal da Bahia\\
   Faculty of Achitecture\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
+  Federal University of Bahia\\
+  Caetano Moura street, 121, \\
+  40210-905, Salvador-BA, Brazil\\
   E-mail: \email{fernando.ribeiro@ufba.br}\\
+  URL: \url{https://arquitetura.ufba.br/}\\
   \emph{and}\\
-  Centro Universitário Senai-Cimatec\\
-
-  % URL: \url{https://www.zeileis.org/}
+  Graduate Program in Modeling in Earth and Environmental Sciences\\
+  State University of Feira de Santana\\
+  Av. Transnordestina\\
+  44036-900, Feira de Santana-BA, Brazil\\
+  URL: \url{https://ppgm.uefs.br/}\\
+  \emph{and}\\
+  Senai-Cimatec University\\
+  Av. Orlando Gomes, 1845,\\
+  41650-010, Salvador-BA, Brazil \\
+  URL: \url{https://www.senaicimatec.com.br/}\\
+  \\
+  Gilney Figueira Zebende\\
+  Graduate Program in Modeling in Earth and Environmental Sciences\\
+  State University of Feira de Santana\\
+  Av. Transnordestina\\
+  44036-900, Feira de Santana-BA, Brazil\\
+  E-mail: \email{gfzebende@uefs.br}\\
+  URL: \url{https://ppgm.uefs.br/}\\
 }
